---
title: "09_20_2017"
author: "John D."
date: "September 18, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

# 22 Introduction

## 22.1 Hypothesis generation vs. hypothesis confirmation

# 23 Model basics

## 23.1 Introduction

### 23.1.1 Prerequisites

```{r}
library(tidyverse)

library(modelr)
options(na.action = na.warn)
```

## 23.2 A simple model

```{r}
ggplot(sim1, aes(x, y)) + 
  geom_point()

models <- tibble(
  a1 = runif(250, -20, 40),
  a2 = runif(250, -5, 5)
)

ggplot(sim1, aes(x, y)) + 
  geom_abline(aes(intercept = a1, slope = a2), data = models, alpha = 1/4) +
  geom_point() 

model1 <- function(a, data) {
  a[1] + data$x * a[2]
}

model1(c(7, 1.5), sim1)

measure_distance <- function(mod, data) {
  diff <- data$y - model1(mod, data)
  sqrt(mean(diff ^ 2))
}
measure_distance(c(7, 1.5), sim1)

sim1_dist <- function(a1, a2) {
  measure_distance(c(a1, a2), sim1)
}

models <- models %>% 
  mutate(dist = purrr::map2_dbl(a1, a2, sim1_dist))
models

ggplot(sim1, aes(x, y)) + 
  geom_point(size = 2, colour = "grey30") + 
  geom_abline(
    aes(intercept = a1, slope = a2, colour = -dist), 
    data = filter(models, rank(dist) <= 10)
  )

ggplot(models, aes(a1, a2)) +
  geom_point(data = filter(models, rank(dist) <= 10), size = 4, colour = "red") +
  geom_point(aes(colour = -dist))

grid <- expand.grid(
  a1 = seq(-5, 20, length = 25),
  a2 = seq(1, 3, length = 25)
  ) %>% 
  mutate(dist = purrr::map2_dbl(a1, a2, sim1_dist))

grid %>% 
  ggplot(aes(a1, a2)) +
  geom_point(data = filter(grid, rank(dist) <= 10), size = 4, colour = "red") +
  geom_point(aes(colour = -dist))


ggplot(sim1, aes(x, y)) + 
  geom_point(size = 2, colour = "grey30") + 
  geom_abline(
    aes(intercept = a1, slope = a2, colour = -dist), 
    data = filter(grid, rank(dist) <= 10)
  )

best <- optim(c(0, 0), measure_distance, data = sim1)
best$par

ggplot(sim1, aes(x, y)) + 
  geom_point(size = 2, colour = "grey30") + 
  geom_abline(intercept = best$par[1], slope = best$par[2])

sim1_mod <- lm(y ~ x, data = sim1)
coef(sim1_mod)
```

### 23.2.1 Exercises

1. One downside of the linear model is that it is sensitive to unusual values because the distance incorporates a squared term. Fit a linear model to the simulated data below, and visualise the results. Rerun a few times to generate different simulated datasets. What do you notice about the model?

```{r}
sim_models <- function(){
  sim1a <- tibble(
  x = rep(1:10, each = 3),
  y = x * 1.5 + 6 + rt(length(x), df = 2)
)

ggplot(sim1a, aes(x, y)) + 
  geom_point(size = 2, colour = "grey30") + 
  geom_abline(intercept = best$par[1], slope = best$par[2])
}

sim_models()
sim_models()
sim_models()
sim_models()
```

The best model does not fit as nicely when you have unusual values. The model is forced to adapt to account for the unusual values. If they were not present, the model would line up more nicely with the rest of the data.

2. One way to make linear models more robust is to use a different distance measure. For example, instead of root-mean-squared distance, you could use mean-absolute distance:

```{r}
measure_distance <- function(mod, data) {
  diff <- data$y - model1(mod, data)
  sqrt(mean(diff ^ 2))
}

measure_distance2 <- function(mod, data) {
  diff <- data$y - model1(mod, data)
  mean(abs(diff))
}

best <- optim(c(0, 0), measure_distance, data = sim1)
best$par

best <- optim(c(0, 0), measure_distance2, data = sim1)
best$par

```

They are similar but not the same.

3. One challenge with performing numerical optimisation is that it’s only guaranteed to find one local optima. What’s the problem with optimising a three parameter model like this?

```{r}
model1 <- function(a, data) {
  a[1] + data$x * a[2] + a[3]
}
```

There are unlimited combinations of a[1] and a[3] that would give you the same optimum y intercept.