---
title: "05_31_2017"
author: "John D."
date: "May 30, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

### 7.6 Patterns and models

```{r}
ggplot(data = faithful) + 
  geom_point(mapping = aes(x = eruptions, y = waiting))

library(modelr)

mod <- lm(log(price) ~ log(carat), data = diamonds)

diamonds2 <- diamonds %>% 
  add_residuals(mod) %>% 
  mutate(resid = exp(resid))

ggplot(data = diamonds2) + 
  geom_point(mapping = aes(x = carat, y = resid))

ggplot(data = diamonds2) + 
  geom_boxplot(mapping = aes(x = cut, y = resid))
```

### 7.7 ggplot2 calls

```{r}
ggplot(data = faithful, mapping = aes(x = eruptions)) + 
  geom_freqpoly(binwidth = 0.25)

ggplot(faithful, aes(eruptions)) + 
  geom_freqpoly(binwidth = 0.25)

diamonds %>% 
  count(cut, clarity) %>% 
  ggplot(aes(clarity, cut, fill = n)) + 
    geom_tile()
```

### 7.8 Learning more

## 8 Workflow: projects

### 8.1 What is real?

### 8.2 Where does your analysis live?

```{r}
getwd()
```

### 8.3 Paths and directories

### 8.4 RStudio projects

```{r}
library(tidyverse)

ggplot(diamonds, aes(carat, price)) + 
  geom_hex()
#ggsave("diamonds.pdf")

#write_csv(diamonds, "diamonds.csv")
```

### 8.5 Summary

##9 Introduction

## 10 Tibbles

### 10.1 Introduction

### 10.2 Creating tibbles

```{r}
iris

as_tibble(iris)

tibble(
  x = 1:5, 
  y = 1, 
  z = x ^ 2 + y
)

tb <- tibble(
  `:)` = "smile", 
  ` ` = "space",
  `2000` = "number"
)
tb

tribble(
  ~x, ~y, ~z,
  #--|--|----
  "a", 2, 3.6,
  "b", 1, 8.5
)
```

### 10.3 Tibbles vs. data.frame

### 10.3.1 Printing


```{r}
tibble(
  a = lubridate::now() + runif(1e3) * 86400,
  b = lubridate::today() + runif(1e3) * 30,
  c = 1:1e3,
  d = runif(1e3),
  e = sample(letters, 1e3, replace = TRUE)
)

nycflights13::flights %>% 
  print(n = 10, width = Inf)

nycflights13::flights %>% 
  View()
```

### 10.3.2 Subsetting

```{r}
df <- tibble(
  x = runif(5),
  y = rnorm(5)
)

df$x
df[["x"]]
df[[1]]

df %>% .$x
df %>% .[["x"]]
```

### 10.4 Interacting with older code

```{r}
#class(as.data.frame(tb))
```

### 10.5 Exercises

1. How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame)

```{r}
print(mtcars)
print(as_tibble(mtcars))
a <- mtcars
b <- as_tibble(mtcars)
class(a)
class(b)
```

It says "tibble" at the top if it is actually a tibble. Also the default number or rows printed in lower with tibbles. The class of a tibble includes "tbl"and "tbl_df".

2. Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?

```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]

tdf <- as_tibble(data.frame(abc = 1, xyz = "a"))
#tdf$x
# ^^ Will not run
tdf[, "xyz"]
tdf[, c("abc", "xyz")]
```

With a dataframe, it will match if the pattern is contained is at the start, but does not match for exact patterns. Can cause you to include more columns than you want when subsetting.

3. If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?

```{r}
tdf
var <- "xyz"
tdf[var]
tdf %>% .[var]
```

4. Practice referring to non-syntactic names in the following data frame by:
    1.  Extracting the variable called `1`.
    2.  Plotting a scatterplot of `1` vs `2`.
    3. Creating a new column called 3 which is 2 divided by 1
    4. Renaming the columns to one, two and three
    
```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
annoying

#1
annoying$`1`
annoying[["1"]]

#2
ggplot(annoying, aes(`1`,`2`)) + geom_point()

#3
annoying$`3` <- annoying$`2` / annoying$`1`
annoying

#4
rename(annoying, one = `1`, two = `2`, three = `3`)
#or
(colnames(annoying) <- c("one","two","three"))
```

5. What does `tibble::enframe()` do? When might you use it?

```{r}
?enframe

As <- c("a","a","A")
Bs <- c("B","b","b")
enframe(c(As,Bs))
enframe(c(a = 5, b = 7))
deframe(df)
```

`enframe()` converts named atomic vectors or lists to two-column data frames. For unnamed vectors, the natural sequence is used as name column. Convienient for creating tibbles quicking using vectors.

6. What option controls how many additional column names are printed at the footer of a tibble?

```{r}
#options(tibble.width = Inf)

# or

# nycflights13::flights %>% 
#  print(n = 10, width = Inf)

```

