---
title: "05_10_2017"
author: "John D."
date: "May 7, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(nycflights13)
library(tidyverse)
```

### 5.2.4 Exercises

1. Find all flights that
    1. Had an arrival delay of two or more hours
    ```{r}
    filter(flights, arr_delay >= 120)
    ```
    2. Flew to Houston (`IAH` or `HOU`)
    ```{r}
    filter(flights, dest == "IAH" | dest == "HOU")
    ```
    3. Were operated by United, American, or Delta
    ```{r}
    filter(flights, carrier == "UA" | carrier == "DL" | carrier == "AA")
    ```
    4. Departed in summer (July, August, and September)
    ```{r}
    filter(flights, month %in% c(7,8,9))
    ```
    5. Arrived more than two hours late, but didn’t leave late
    ```{r}
    filter(flights, arr_delay > 120 & dep_delay <= 0)
    ```
    6. Were delayed by at least an hour, but made up over 30 minutes in flight
    ```{r}
    filter(flights, dep_delay >= 60 & (dep_delay - arr_delay) > 30)
    ```
    7. Departed between midnight and 6am (inclusive)
    ```{r}
    filter(flights, dep_time %in% 0:600)
    ```

2. Another useful dplyr filtering helper is `between()`. What does it do? Can you use it to simplify the code needed to answer the previous challenges?

  This is a shortcut for x >= left & x <= right. Can use it for finding flights between certain months and times
  
3. How many flights have a missing `dep_time`? What other variables are missing? What might these rows represent?

```{r}
filter(flights, is.na(dep_time))
```

  Delays and actual arrival and flight times aren't listed. These flights were most likely cancelled and never actually took place.
  
4. Why is `NA ^ 0` not missing? Why is `NA | TRUE` not missing? Why is `FALSE & NA` not missing? Can you figure out the general rule? (`NA * 0` is a tricky counterexample!)

They are not missing because they have will have a default value no matter what the NA can be. 1 or -1 , TRUE, FALSE. The last one does not work because we cannot multiple nothing by 0. There is no default value.

### 5.3.1 Exercises

1. How could you use `arrange()` to sort all missing values to the start? (Hint: use `is.na()`).

```{r}
arrange(flights, desc(is.na(dep_time)))
```

2. Sort `flights` to find the most delayed flights. Find the flights that left earliest

```{r}
arrange(flights, desc(dep_delay))
arrange(flights, dep_delay)

```

3. Sort `flights` to find the fastest flights.

```{r}
arrange(flights, air_time)
```

4. Which flights travelled the longest? Which travelled the shortest?

```{r}
head(arrange(flights, desc(distance)))
head(arrange(flights, distance))
```

### 5.4.1 Exercises

1. Brainstorm as many ways as possible to select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from flights.

```{r}
head(select(flights, starts_with("dep"), starts_with("arr")),1)
head(select(flights, dep_time, dep_delay, arr_time, arr_delay),1)
head(select(flights, dep_time:arr_delay, -starts_with("sched")),1)
```

2. What happens if you include the name of a variable multiple times in a `select()` call?

```{r}
head(select(flights, dep_time, dep_time, dep_time), 1)
```

Only get the variable once

3. What does the `one_of()` function do? Why might it be helpful in conjunction with this vector?

```{r}
?one_of
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
head(select(flights, one_of(vars)),1)
```

Allows you to select columns by passing a character vector to the function

4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

```{r}
select(flights, contains("TIME"))
```

Nope, it matches while being case insensitive. Can turn off by setting `ignore.case = F`

### 5.5.2 Exercises

1. Currently `dep_time` and `sched_dep_time` are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}
mutate(flights,
       dep_time_mins = dep_time %/% 100 * 60 + dep_time %% 100,
       sched_dep_time_mins = sched_dep_time %/% 100 * 60 + sched_dep_time %% 100) %>%
  select(dep_time, dep_time_mins, sched_dep_time, sched_dep_time_mins)

```

2. Compare `air_time` with `arr_time - dep_time`. What do you expect to see? What do you see? What do you need to do to fix it?

```{r}
mutate(flights,
       air_time2 = arr_time - dep_time,
       air_time_diff = air_time2 - air_time) %>%
  filter(air_time_diff != 0) %>%
  select(air_time, air_time2, dep_time, arr_time, dest)
```

`dep_time` and `arv_time` are dependent on the airport's location and are thereby affected by timezones.

3. Compare `dep_time`, `sched_dep_time`, and `dep_delay`. How would you expect those three numbers to be related?

`dep_delay` is the result of subtracting `sched_dep_time` from `dep_time`

4. Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for `min_rank()`.

```{r}
mutate(flights,
       dep_delay_rank = min_rank(-dep_delay)) %>%
  arrange(dep_delay_rank) %>% 
  filter(dep_delay_rank <= 10)
```

5. What does `1:3` + `1:10` return? Why?

```{r}
1:3 + 1:10
```

This vector is returned because it will apply the operation until every element in the longest vector has been used while recycling the elements of the smaller vector.

6. What trigonometric functions does R provide?

```{r}
?Trig
```
cos(x), sin(x), tan(x), acos(x), asin(x), atan(x), atan2(y, x), cospi(x), sinpi(x), and (tanpi(x))
