---
title: "Untitled"
author: "John D."
date: "August 23, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(error = TRUE, echo = TRUE)
```

# 19 Functions

## 19.5 Function arguments

```{r}
# Compute confidence interval around mean using normal approximation
mean_ci <- function(x, conf = 0.95) {
  se <- sd(x) / sqrt(length(x))
  alpha <- 1 - conf
  mean(x) + se * qnorm(c(alpha / 2, 1 - alpha / 2))
}

x <- runif(100)
mean_ci(x)

mean_ci(x, conf = 0.99)

# Good
mean(1:10, na.rm = TRUE)

# Bad
mean(x = 1:10, , FALSE)
mean(, TRUE, x = c(1:10, NA))

# Good
average <- mean(feet / 12 + inches, na.rm = TRUE)

# Bad
average<-mean(feet/12+inches,na.rm=TRUE)
```

### 19.5.1 Choosing names

### 19.5.2 Checking values

```{r}
wt_mean <- function(x, w) {
  sum(x * w) / sum(w)
}
wt_var <- function(x, w) {
  mu <- wt_mean(x, w)
  sum(w * (x - mu) ^ 2) / sum(w)
}
wt_sd <- function(x, w) {
  sqrt(wt_var(x, w))
}

wt_mean(1:6, 1:3)

wt_mean <- function(x, w) {
  if (length(x) != length(w)) {
    stop("`x` and `w` must be the same length", call. = FALSE)
  }
  sum(w * x) / sum(w)
}

wt_mean(1:6, 1:3)

wt_mean <- function(x, w, na.rm = FALSE) {
  if (!is.logical(na.rm)) {
    stop("`na.rm` must be logical")
  }
  if (length(na.rm) != 1) {
    stop("`na.rm` must be length 1")
  }
  if (length(x) != length(w)) {
    stop("`x` and `w` must be the same length", call. = FALSE)
  }
  
  if (na.rm) {
    miss <- is.na(x) | is.na(w)
    x <- x[!miss]
    w <- w[!miss]
  }
  sum(w * x) / sum(w)
}

wt_mean <- function(x, w, na.rm = FALSE) {
  stopifnot(is.logical(na.rm), length(na.rm) == 1)
  stopifnot(length(x) == length(w))
  
  if (na.rm) {
    miss <- is.na(x) | is.na(w)
    x <- x[!miss]
    w <- w[!miss]
  }
  sum(w * x) / sum(w)
}
wt_mean(1:6, 6:1, na.rm = "foo")
```

### 19.5.3 Dot-dot-dot (…)

```{r}
commas <- function(...) stringr::str_c(..., collapse = ", ")
commas(letters[1:10])

rule <- function(..., pad = "-") {
  title <- paste0(...)
  width <- getOption("width") - nchar(title) - 5
  cat(title, " ", stringr::str_dup(pad, width), "\n", sep = "")
}
rule("Important output")

x <- c(1, 2)
sum(x, na.mr = TRUE)
```

### 19.5.4 Lazy evaluation

### 19.5.5 Exercises

1. What does `commas(letters, collapse = "-")` do? Why?

```{r}
commas(letters, collapse = "-")
```

Throws an error in my case

2. It’d be nice if you could supply multiple characters to the `pad` argument, e.g. `rule("Title", pad = "-+")`. Why doesn’t this currently work? How could you fix it?

```{r}
rule("Title", pad = "-+")

rule2 <- function(..., pad = "-") {
  title <- paste0(...)
  width <- (getOption("width") - nchar(title) - 5)/nchar(pad)
  cat(title, " ", stringr::str_dup(pad, width), "\n", sep = "")
}
rule2("Title", pad = "-+")

```

The width does not take the length of pad is not taken into account, as a result there is too much padding. Need to scale width variable to the length of the padding argument.

3. What does the `trim` argument to `mean()` do? When might you use it?

```{r}
mean(c(1.543,12.53453,1.00001,214,6546))
mean(c(1.543,12.53453,1.00001,214,6546), trim = 0)
mean(c(1.543,12.53453,1.00001,214,6546), trim = .25)
mean(c(1.543,12.53453,1.00001,214,6546), trim = .5)

?mean
```

the fraction (0 to 0.5) of observations to be trimmed from each end of x before the mean is computed. Values of trim outside that range are taken as the nearest endpoint.

4. The default value for the method argument to `cor()` is `c("pearson", "kendall", "spearman")`. What does that mean? What value is used by default?

```{r}
?cor
```

The character string is the available options for the argument `method`. By default `pearson` is used.

## 19.6 Return values

### 19.6.1 Explicit return statements

```{r}
complicated_function <- function(x, y, z) {
  if (length(x) == 0 || length(y) == 0) {
    return(0)
  }
    
  # Complicated code here
}

f <- function() {
  if (x) {
    # Do 
    # something
    # that
    # takes
    # many
    # lines
    # to
    # express
  } else {
    # return something short
  }
}

f <- function() {
  if (!x) {
    return(something_short)
  }

  # Do 
  # something
  # that
  # takes
  # many
  # lines
  # to
  # express
}
```

### 19.6.2 Writing pipeable functions

```{r}
show_missings <- function(df) {
  n <- sum(is.na(df))
  cat("Missing values: ", n, "\n", sep = "")
  
  invisible(df)
}

show_missings(mtcars)

x <- show_missings(mtcars) 
class(x)
dim(x)
mtcars %>% 
  show_missings() %>% 
  mutate(mpg = ifelse(mpg < 20, NA, mpg)) %>% 
  show_missings() 
```

## 19.7 Environment

```{r}
f <- function(x) {
  x + y
} 

y <- 100
f(10)

y <- 1000
f(10)

`+` <- function(x, y) {
  if (runif(1) < 0.1) {
    sum(x, y)
  } else {
    sum(x, y) * 1.1
  }
}
table(replicate(1000, 1 + 2))

rm(`+`)
```

# 20 Vectors

## 20.1 Introduction

### 20.1.1 Prerequisites

```{r}
library(tidyverse)
```

## 20.2 Vector basics

```{r}
typeof(letters)
typeof(1:10)

x <- list("a", "b", 1:10)
length(x)
```

## 20.3 Important types of atomic vector

### 20.3.1 Logical

```{r}
1:10 %% 3 == 0

c(TRUE, TRUE, FALSE, NA)
```

### 20.3.2 Numeric

```{r}
typeof(1)

typeof(1L)

1.5L

x <- sqrt(2) ^ 2
x

x - 2

c(-1, 0, 1) / 0
```

### 20.3.3 Character

```{r}
x <- "This is a reasonably long string."
pryr::object_size(x)

y <- rep(x, 1000)
pryr::object_size(y)
```

### 20.3.4 Missing values

```{r}
NA
NA_integer_
NA_real_
NA_character_
```

### 20.3.5 Exercises

1. Describe the difference between `is.finite(x)` and `!is.infinite(x)`

```{r}
x <- Inf
y <- 123
z <- NA

is.finite(x)
!is.infinite(x)
is.finite(y)
!is.infinite(y)
is.finite(z)
!is.infinite(z)
```

The difference appears when handling `NA`. for `is.finite()`, `NA` is not finite, and for `is.infinite()`, `NA` is not infinite.

2. Read the source code for `dplyr::near()` (Hint: to see the source code, drop the `()`). How does it work?

```{r}
dplyr::near

# function (x, y, tol = .Machine$double.eps^0.5) 
# {
#     abs(x - y) < tol
# }

.Machine$double.eps^0.5

```

Uses predetermined tolerance level (1.490116e-08) and decides if deference between the two values is below that threshold.

3. A logical vector can take 3 possible values. How many possible values can an integer vector take? How many possible values can a double take? Use google to do some research.

Using 32-bit R, integer vectors can contain -2\*10^9 through 2\*10^9 plus `NA`. Doubles work with a precision of 53 bits and can contain 2e-308 to 2e+308 along with `NaN`, +0, -0, -Inf, Inf, and `NA`.

4. Brainstorm at least four functions that allow you to convert a double to an integer. How do they differ? Be precise.

```{r}
x <- 234.643
typeof(x)

y <- integer(x)
typeof(y)

y <- as.integer(x)
typeof(y)

y <- as.factor(x)
typeof(y)

y <- factor(x)
typeof(y)

y <- as.vector(x, mode = "integer")
typeof(y)

y <- round(x, 0)
typeof(y) # still double
```

5. What functions from the readr package allow you to turn a string into logical, integer, and double vector?

```{r}
library(readr)
string <- c("0", "Hello", "123", "TRUE", "NA", "1")

(a <- readr::parse_logical(string))
typeof(a)

(a <- readr::parse_integer(string))
typeof(a)

(a <- readr::parse_double(string))
typeof(a)
```