---
title: "07_19_2017"
author: "John D."
date: "July 14, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 14.4 Tools

## 14.4.1 Detect matches

```{r}
library(tidyverse)
library(stringr)
x <- c("apple", "banana", "pear")
str_detect(x, "e")

sum(str_detect(words, "^t"))

mean(str_detect(words, "[aeiou]$"))

no_vowels_1 <- !str_detect(words, "[aeiou]")

no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
identical(no_vowels_1, no_vowels_2)

words[str_detect(words, "x$")]

str_subset(words, "x$")

df <- tibble(
  word = words, 
  i = seq_along(word)
)
df %>% 
  filter(str_detect(words, "x$"))

x <- c("apple", "banana", "pear")
str_count(x, "a")

mean(str_count(words, "[aeiou]"))

df %>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]")
  )

str_count("abababa", "aba")

#str_view_all("abababa", "aba")
```

### 14.4.2 Exercises

1. For each of the following challenges, try solving it by using both a single regular expression, and a combination of multiple `str_detect()` calls.

    1. Find all words that start or end with x.
    
    ```{r}
words[str_detect(words, "^x|x$")]
words[str_detect(words,"^x") | str_detect(words,"x$")]
```

    2. Find all words that start with a vowel and end with a consonant
    
    ```{r}
words[str_detect(words, "^[aeiou].*[^aeiou]$")]
words[str_detect(words, "^[aeiou]") & str_detect(words, "[^aeiou]$")]
```

    3. Are there any words that contain at least one of each different vowel?
    
    ```{r}
### Test  
test <- c("allotetraploid","autotetraploidy")
test[str_detect(sapply(test, function(x) paste(sort(unlist(str_split(x,""))), collapse = "")),"a{1,}.*e{1,}.*i{1,}.*o{1,}.*u")]

### Actual 
words[str_detect(sapply(words, function(x) paste(sort(unlist(str_split(x,""))), collapse = "")),"a{1,}.*e{1,}.*i{1,}.*o{1,}.*u")]
words[str_detect(words,"a")&str_detect(words,"e")&str_detect(words,"i")&str_detect(words,"o")&str_detect(words,"u")]
```

2. What word has the highest number of vowels? What word has the highest proportion of vowels? (Hint: what is the denominator?)

```{r}
words[which(str_count(words,"[aeiou]") == max(str_count(words,"[aeiou]")))]
```

    A is the word with the highest proportion of vowels in the word object.
