---
title: "07_19_2017"
author: "John D."
date: "July 14, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 14.4 Tools

## 14.4.1 Detect matches

```{r}
library(tidyverse)
library(stringr)
x <- c("apple", "banana", "pear")
str_detect(x, "e")

sum(str_detect(words, "^t"))

mean(str_detect(words, "[aeiou]$"))

no_vowels_1 <- !str_detect(words, "[aeiou]")

no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
identical(no_vowels_1, no_vowels_2)

words[str_detect(words, "x$")]

str_subset(words, "x$")

df <- tibble(
  word = words, 
  i = seq_along(word)
)
df %>% 
  filter(str_detect(words, "x$"))

x <- c("apple", "banana", "pear")
str_count(x, "a")

mean(str_count(words, "[aeiou]"))

df %>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]")
  )

str_count("abababa", "aba")

#str_view_all("abababa", "aba")
```

### 14.4.2 Exercises

1. For each of the following challenges, try solving it by using both a single regular expression, and a combination of multiple `str_detect()` calls.

    1. Find all words that start or end with x.
    
    ```{r}
words[str_detect(words, "^x|x$")]
words[str_detect(words,"^x") | str_detect(words,"x$")]
```

    2. Find all words that start with a vowel and end with a consonant
    
    ```{r}
words[str_detect(words, "^[aeiou].*[^aeiou]$")]
words[str_detect(words, "^[aeiou]") & str_detect(words, "[^aeiou]$")]
```

    3. Are there any words that contain at least one of each different vowel?
    
    ```{r}
### Test  
test <- c("allotetraploid","autotetraploidy")
test[str_detect(sapply(test, function(x) paste(sort(unlist(str_split(x,""))), collapse = "")),"a{1,}.*e{1,}.*i{1,}.*o{1,}.*u")]

### Actual 
words[str_detect(sapply(words, function(x) paste(sort(unlist(str_split(x,""))), collapse = "")),"a{1,}.*e{1,}.*i{1,}.*o{1,}.*u")]
words[str_detect(words,"a")&str_detect(words,"e")&str_detect(words,"i")&str_detect(words,"o")&str_detect(words,"u")]
```

2. What word has the highest number of vowels? What word has the highest proportion of vowels? (Hint: what is the denominator?)

```{r}
vowel_wds <- words[which(str_count(words,"[aeiou]") == max(str_count(words,"[aeiou]")))]
vowel_wds
vowel_wds[which(str_count(vowel_wds,"[aeiou]")/str_count(vowel_wds,"[a-z]") == max(str_count(vowel_wds,"[aeiou]")/str_count(vowel_wds,"[a-z]")))]

```

    A is the shortest word with the highest proportion of vowels in the word object.

## 14.4.3 Extract matches

```{r}
length(sentences)
head(sentences)

colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
colour_match

has_colour <- str_subset(sentences, colour_match)
matches <- str_extract(has_colour, colour_match)
head(matches)

more <- sentences[str_count(sentences, colour_match) > 1]
str_view_all(more, colour_match)

str_extract(more, colour_match)
str_extract_all(more, colour_match)

str_extract_all(more, colour_match, simplify = TRUE)

x <- c("a", "a b", "a b c")
str_extract_all(x, "[a-z]", simplify = TRUE)
```

### 14.4.3.1 Exercises

1.In the previous example, you might have noticed that the regular expression matched “flickered”, which is not a colour. Modify the regex to fix the problem

```{r}
more <- sentences[str_count(sentences, colour_match) > 1]
colours <- c("\\b(red", "orange", "yellow", "green", "blue", "purple)\\b")
colour_match <- str_c(colours, collapse = "|")
colour_match
str_extract_all(more, colour_match, simplify = TRUE)
```

2. From the Harvard sentences data, extract:
    1. The first word from each sentence
    
    ```{r}
first <- "^\\w*"
str_extract(sentences, first)
```

    2. All words ending in `ing`.
    
    ```{r}
ings <- "[a-zA-Z]*ing\\b"
str_extract_all(sentences, ings, simplify = T) %>%
  as.data.frame() %>%
  filter(V1 != "") %>%
  unlist() %>%
  unique() %>%
  as.character()
```

    3. All plurals
    
    ```{r}
plurals <- "\\b[a-zA-Z]{3,}s\\b"
str_extract_all(sentences, plurals, simplify = T) %>%
  as.data.frame() %>%
  filter(V1 != "") %>%
  unlist() %>%
  unique() %>%
  as.character()
```

## 10.4.4 Grouped Matches

```{r}
noun <- "(a|the) ([^ ]+)"
has_noun <- sentences %>%
  str_subset(noun) %>%
  head(10)
has_noun %>%
  str_extract(noun)

has_noun %>%
  str_match(noun)

tibble(sentence = sentences) %>%
  tidyr::extract(
    sentence, c("article", "noun"), "(a|the) ([^ ]+)",
    remove = FALSE
  )
```

### 10.4.4.1 Exercises

1. Find all words that come after a “number” like “one”, “two”, “three” etc. Pull out both the number and the word

```{r}
nums <- "\\b(one|two|three|four|five|six|seven|eight|nine|ten)\\b ([^ ]+)"
tibble(sentence = sentences) %>%
  tidyr::extract(
    sentence, c("number", "word"), nums,
    remove = FALSE
  ) %>%
  filter(!is.na(number))

## or

sentences[str_detect(sentences, nums)] %>%
  str_extract_all(nums) %>%
  unlist()
```

2. Find all contractions. Separate out the pieces before and after the apostrophe.

```{r}
contracts <- "([^ ]+)'([^ ])+"
sentences[str_detect(sentences, contracts)] %>%
  str_extract_all(contracts) %>%
  unlist()
```

## 14.4.5 Replacing matches

```{r}
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-")
str_replace_all(x, "[aeiou]", "-")

x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))

sentences %>% 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>% 
  head(5)

```

### 14.4.5.1 Exercises

1. Replace all forward slashes in a string with backslashes.
```{r}
test <- "a/b/c/d//e"
writeLines(test)
writeLines(str_replace_all(test, "/","\\\\"))
```

2. Implement a simple version of str_to_lower() using replace_all()

```{r}
test <- "LOUD NOISES"
str_replace_all(test, c("L" = "l", "O" = "o", "U" = "u", "D" = "d", "N" = "n", "I" = "i", "S" = "s", "E" = "e"))
```

3. Switch the first and last letters in words. Which of those strings are still words?

```{r}
words[which(str_replace_all(words, "(^\\w)(.*)(\\w$)","\\3\\2\\1") == words)]
```

## 14.4.6 Splitting

```{r}
sentences %>%
  head(5) %>% 
  str_split(" ")

"a|b|c|d" %>% 
  str_split("\\|") %>% 
  .[[1]]

sentences %>%
  head(5) %>% 
  str_split(" ", simplify = TRUE)

fields <- c("Name: Hadley", "Country: NZ", "Age: 35")
fields %>% str_split(": ", n = 2, simplify = TRUE)

x <- "This is a sentence.  This is another sentence."
str_view_all(x, boundary("word"))

str_split(x, " ")[[1]]
str_split(x, boundary("word"))[[1]]
```

### 14.4.6.1 Exercises

1. Split up a string like `"apples, pears, and bananas"` into individual components.

```{r}
x <- "apples, pears, and bananas" 
str_split(x, boundary("word"))[[1]]
```

2. Why is it better to split up by `boundary("word")` than `" "`?

Multiple spaces in a row will lead to empty strings

3. What does splitting with an empty string ("") do? Experiment, and then read the documentation.

```{r}
str_split(x, "")
```

Splits every character in the string into a new string

# 14.5 Other types of pattern

```{r}
str_view(fruit, "nana")
str_view(fruit, regex("nana"))

bananas <- c("banana", "Banana", "BANANA")
str_view(bananas, "banana")

str_view(bananas, regex("banana", ignore_case = TRUE))

x <- "Line 1\nLine 2\nLine 3"
str_extract_all(x, "^Line")[[1]]

str_extract_all(x, regex("^Line", multiline = TRUE))[[1]]

phone <- regex("
  \\(?     # optional opening parens
  (\\d{3}) # area code
  [)- ]?   # optional closing parens, dash, or space
  (\\d{3}) # another three numbers
  [ -]?    # optional space or dash
  (\\d{3}) # three more numbers
  ", comments = TRUE)

str_match("514-791-8141", phone)

microbenchmark::microbenchmark(
  fixed = str_detect(sentences, fixed("the")),
  regex = str_detect(sentences, "the"),
  times = 20
)

a1 <- "\u00e1"
a2 <- "a\u0301"
c(a1, a2)

a1 == a2

str_detect(a1, fixed(a2))
str_detect(a1, coll(a2))

i <- c("I", "İ", "i", "ı")
i

str_subset(i, coll("i", ignore_case = TRUE))

str_subset(i, coll("i", ignore_case = TRUE, locale = "tr"))

stringi::stri_locale_info()

x <- "This is a sentence."
str_view_all(x, boundary("word"))

str_extract_all(x, boundary("word"))
```

### 14.5.1 Exercises

1. How would you find all strings containing `\` with `regex()` vs. with `fixed()`?

```{r}
x <- c("a\\b\\c", "test")
str_detect(x, "\\\\")
str_detect(x, fixed("\\"))
```

2. What are the five most common words in `sentences`?

```{r}
str_extract_all(sentences, boundary("word")) %>%
  unlist() %>%
  str_to_lower() %>%
  tibble() %>%
  set_names("word") %>%
  group_by(word) %>%
  count(sort = TRUE) %>%
  head(5)
```

# 14.6 Other uses of regular expressions

```{r}
apropos("replace")
head(dir(pattern = "\\.Rmd$"))
```

# 14.7 stringi

### 14.7.1 Exercises

1.Find the stringi functions that

  1. Count the number of words  
    `stri_count_words()`  
  2. Find duplicated strings  
    `stri_duplicated()`  
  3. Generate random text.  
    `stri_rand_strings()`
    
2. How do you control the language that stri_sort() uses for sorting?

Using the `local=""` argument