---
title: "06_07_2017"
author: "John D."
date: "June 5, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

# 11 Data import

## 11.2 Getting started

```{r}
#heights <- read_csv("data/heights.csv")

read_csv("a,b,c
1,2,3
4,5,6")

read_csv("The first line of metadata
  The second line of metadata
  x,y,z
  1,2,3", skip = 2)

read_csv("# A comment I want to skip
  x,y,z
  1,2,3", comment = "#")

read_csv("1,2,3\n4,5,6", col_names = FALSE)

read_csv("1,2,3\n4,5,6", col_names = c("x", "y", "z"))

read_csv("a,b,c\n1,2,.", na = ".")
```

### 11.2.1 Compared to base R

#### 11.2.2 Exercises

1. What function would you use to read a file where fields were separated with “|”?

```{r}
read_delim("fruits.txt", "|")
```

2. Apart from `file`, `skip`, and `comment`, what other arguments do `read_csv()` and `read_tsv()` have in common?

`read_csv()` and `read_tsv()` also have `col_names`, `col_types`, `locale`, `na`, `quoted_na`, `quote`, `trim_ws`, `n_max`, `guess_max`, and `progress` in common.

3. What are the most important arguments to `read_fwf()`?

The most important arguments are `file` and `col_positions`. `col_positions` is created using `fwf_widths()`, `fwf_empty()`, and `fwf_positions()`

4. Sometimes strings in a CSV file contain commas. To prevent them from causing problems they need to be surrounded by a quoting character, like `"` or `'`. By convention, `read_csv()` assumes that the quoting character will be `"`, and if you want to change it you’ll need to use `read_delim()` instead. What arguments do you need to specify to read the following text into a data frame?

```{r}
read_delim("x,y\n1,'a,b'", ",", quote = "'")
```

5. Identify what is wrong with each of the following inline CSV files. What happens when you run the code?

```{r}
read_csv("a,b\n1,2,3\n4,5,6")
```

2 column names, but the all rows list 3 elements instead of 2.

```{r}
read_csv("a,b,c\n1,2\n1,2,3,4")
```

Header says theres 3 columns. 1st row is 1 element too short and 2nd row is 1 element too long

```{r}
read_csv("a,b\n\"1")
```

Header says theres 2 columns. 1st row is 1 element too short. Also theres an extra double quote, not sure if it's missing a mate or supposed to be its own element.

```{r}
read_csv("a,b\n1,2\na,b")
```

Technically nothing wrong, it reads exactly as expected. Not sure what person was trying to do. Maybe wanted 1 and 2 to be the headers but switched them.

```{r}
read_csv("a;b\n1;3")
read_delim("a;b\n1;3", ";")
```

Assuming here they wanted `;` to be the delimiter.

## 11.3 Parsing a vector

```{r}
str(parse_logical(c("TRUE", "FALSE", "NA")))
str(parse_integer(c("1", "2", "3")))
str(parse_date(c("2010-01-01", "1979-10-14")))
parse_integer(c("1", "231", ".", "456"), na = ".")
x <- parse_integer(c("123", "345", "abc", "123.45"))
x
problems(x)
```

### 11.3.1 Numbers

```{r}
parse_double("1.23")
parse_double("1,23", locale = locale(decimal_mark = ","))
parse_number("$100")
parse_number("20%")
parse_number("It cost $123.45")
parse_number("$123,456,789")
parse_number("123.456.789", locale = locale(grouping_mark = "."))
parse_number("123'456'789", locale = locale(grouping_mark = "'"))
```

### 11.3.2 Strings

```{r}
charToRaw("John")

x1 <- "El Ni\xf1o was particularly bad this year"
x2 <- "\x82\xb1\x82\xf1\x82\xc9\x82\xbf\x82\xcd"

x1
x2

parse_character(x1, locale = locale(encoding = "Latin1"))
parse_character(x2, locale = locale(encoding = "Shift-JIS"))

guess_encoding(charToRaw(x1))
guess_encoding(charToRaw(x2))
```

### 11.3.3 Factors

```{r}
fruit <- c("apple", "banana")
parse_factor(c("apple", "banana", "bananana"), levels = fruit)
```

### 11.3.4 Dates, date-times, and times

```{r}
parse_datetime("2010-10-01T2010")
parse_datetime("20101010")
parse_date("2010-10-01")
library(hms)
parse_time("01:10 am")
parse_time("20:10:01")

parse_date("01/02/15", "%m/%d/%y")
parse_date("01/02/15", "%d/%m/%y")
parse_date("01/02/15", "%y/%m/%d")
parse_date("1 janvier 2015", "%d %B %Y", locale = locale("fr"))
```

#### 11.3.5 Exercises

1. What are the most important arguments to locale()?

`encoding()` and `decimal_mark` are the most important arguments in my opinion.

2. What happens if you try and set `decimal_mark` and `grouping_mark` to the same character? What happens to the default value of `grouping_mark` when you set `decimal_mark` to “,”? What happens to the default value of `decimal_mark` when you set the `grouping_mark` to “.”?

```{r}
#parse_number("123,456,789.0", locale = locale(grouping_mark = ".", decimal_mark = "."))
parse_number("123,456,789.0", locale = locale(decimal_mark = ","))
parse_number("123,456,789.0", locale = locale(grouping_mark = "."))

```

The two must both be different. When `decimal_mark = ","`, `grouping_mark = "."`. When `grouping_mark = "."`, `decimal_mark = ","` 
3. I didn’t discuss the `date_format` and `time_format` options to locale(). What do they do? Construct an example that shows when they might be useful.

```{r}
?parse_date

parse_datetime("2010/01/01 12:00", "%Y/%m/%d %H:%M",
  locale = locale(tz = "US/Central"))
parse_datetime("2010/01/01 12:00", "%Y/%m/%d %H:%M",
  locale = locale(tz = "US/Eastern"))

parse_date("1 janvier 2015", "%d %B %Y")
parse_date("1 janvier 2015", "%d %B %Y", locale = locale("fr"))
```

Useful for listing the timezones for your dates and for translating different dates.

4. If you live outside the US, create a new locale object that encapsulates the settings for the types of file you read most commonly.

```{r}
locale <- locale("de")
parse_date("10 oktober 1993", "%d %B %Y", locale = locale)
```

5.What’s the difference between `read_csv()` and `read_csv2()`

`read_csv()` uses `,` as the delimiter while `read_csv2()` uses `;` as the delimiter. `;` is used as the delimiter because it is common to use `,` as the decimal separator in European countries.

6. What are the most common encodings used in Europe? What are the most common encodings used in Asia? Do some googling to find out.

Europe: ISO 8859-1 and ISO-LATIN-1
Asia: Shift-JIS(Japan), GB18030(China), EUC-KR(Korea)

7. Generate the correct format string to parse each of the following dates and times:

```{r}
d1 <- "January 1, 2010"
parse_date(d1, "%B %d, %Y")

d2 <- "2015-Mar-07"
parse_date(d2, "%Y-%b-%d")

d3 <- "06-Jun-2017"
parse_date(d3, "%d-%b-%Y")

d4 <- c("August 19 (2015)", "July 1 (2015)")
parse_date(d4, "%B %d %.%Y%.")

d5 <- "12/30/14" # Dec 30, 2014
parse_date(d5, "%m/%d/%y")

t1 <- "1705"
parse_time(t1, "%H%M")

t2 <- "11:15:10.12 PM"
parse_time(t2, "%I:%M:%OS %p")
```