---
title: "07_12_2017"
author: "John D."
date: "July 10, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 14 Strings

## 14.1 Introduction

### 14.1.1 Prerequisites

```{r}
library(tidyverse)
library(stringr)
```


## 14.2 String basics

```{r}
string1 <- "This is a string"
string2 <- 'If I want to include a "quote" inside a string, I use single quotes'

double_quote <- "\"" # or '"'
single_quote <- '\'' # or "'"

x <- c("\"", "\\")
x
writeLines(x)

x <- "\u00b5"
x

c("one", "two", "three")
```


### 14.2.1 String length

```{r}
str_length(c("a", "R for data science", NA))
```

### 14.2.2 Combining strings

```{r}
str_c("x", "y")
str_c("x", "y", "z")

str_c("x", "y", sep = ", ")

x <- c("abc", NA)
str_c("|-", x, "-|")

str_c("|-", str_replace_na(x), "-|")

str_c("prefix-", c("a", "b", "c"), "-suffix")

name <- "Hadley"
time_of_day <- "morning"
birthday <- FALSE

str_c(
  "Good ", time_of_day, " ", name,
  if (birthday) " and HAPPY BIRTHDAY",
  "."
)

str_c(c("x", "y", "z"), collapse = ", ")
```

### 14.2.3 Subsetting strings

```{r}
x <- c("Apple", "Banana", "Pear")
str_sub(x, 1, 3)

str_sub(x, -3, -1)

str_sub("a", 1, 5)

str_sub(x, 1, 1) <- str_to_lower(str_sub(x, 1, 1))
x
```

### 14.2.4 Locales

```{r}
str_to_upper(c("i", "ı"))

str_to_upper(c("i", "ı"), locale = "tr")

x <- c("apple", "eggplant", "banana")

str_sort(x, locale = "en")

str_sort(x, locale = "haw")

```

### 14.2.5 Exercises

1. In code that doesn’t use stringr, you’ll often see `paste()` and `paste0()`. What’s the difference between the two functions? What stringr function are they equivalent to? How do the functions differ in their handling of `NA`?

`paste()` will concatenate the objects together and by default separate them by a space, while `paste0()` will concatenate the objects together with no separator. `paste()` == `str_c(x,y, sep = " ")` and `paste0()` == `str_c()`. `paste()` and `paste0()` will keep the NA, but `str_c` will proprogate the NA and everything concatenated to it collectively becomes NA.

2. In your own words, describe the difference between the `sep` and `collapse` arguments to `str_c()`.

`sep` is what string to put between each argument when creating new strings from several strings.  While `collapse` is what to put between each string when collapsing many strings into one. `sep` is what goes between the two arguments when concatenating and `collapse` is what goes between each of the new created strings when collapsing the new strings into one combined string.

```{r}
str_c(c("a","b","c"), c("e","f","g"), sep = "_")
str_c(c("a","b","c"), c("e","f","g"), collapse = "-")
str_c(c("a","b","c"), c("e","f","g"), sep = "_", collapse = "-")
```

3. Use str_length() and str_sub() to extract the middle character from a string. What will you do if the string has an even number of characters?

```{r}
x <- c("telephone", "computer")
str_length(x)
str_sub(x, ceiling(str_length(x)/2), ceiling(str_length(x)/2))
```

If it has an even number of characters, the first of the 2 numbers in the middle will be selected.

4. What does `str_wrap()` do? When might you want to use it?

Wrap strings into nicely formatted paragraphs. Use when you want to make an easier to read text output with a fixed width

```{r}
thanks_path <- file.path(R.home("doc"), "THANKS")
thanks <- str_c(readLines(thanks_path), collapse = "\n")
thanks <- word(thanks, 1, 3, fixed("\n\n"))
cat(str_wrap(thanks), "\n")
```

5. What does `str_trim()` do? What’s the opposite of `str_trim()`?

Trim whitespace from start and end of string. the opposite would be `str_pad()`

```{r}
x <- "  word  "
x
x <- str_trim(x)
x
x <- str_pad(x, 8, "both")
x
```

6. Write a function that turns (e.g.) a vector `c("a", "b", "c")` into the string `a, b, and c. Think carefully about what it should do if given a vector of length 0, 1, or 2.

```{r}
fake_func <- function(x) {
  if(length(x) == 0) print("String is empty")
  if(length(x) == 1) return(x)
  if(length(x) == 2){
    return(str_c(x, collapse = " and "))
  }
  if(length(x) >= 3){
    new_head <- paste(x[1:length(x)-1], collapse = ", ")
    new_tail <- paste0(" and ", x[length(x)])
    new <- paste(new_head,new_tail, sep = ",")
    return(new)
  }
}
w <- c()
x <- c("one")
y <- c("one", "two")
z <- c("one","two","three","four")

fake_func(w)
fake_func(x)
fake_func(y)
fake_func(z)
```

## Regexone tutorial

### Part one

1. "abc"
2. "123"
3. ".*\."
4. "[cmf]an"
5. "[^b]og"
6. "[A-C][nop][a-c]"
7. "waz{2,}up"
8. "aab*c+"
9. "\d+ files? found\?"
10. "\d\.\s+abc"
11. "^M.*l$"
12. "^(file.*)+\.pdf$"
13. "(\w+\s+(\d+))"
14. "(\d+)x(\d+)"
15. "I love (cats|dogs)"
16. ".*"

### Decimals

1. ".*[^p]$"
2. "1?\s?\(?(\d{3})\)?[\s-]?\d{3}[\s-]?\d{4}"
3. "([^\+@]+).*"
4. "<((a|div)).*"
5. "(\w+)\.((jpg|png|gif))$"
6. "^\s*(.*)\s*$"
7. ".*widget.List.(\w+)\(([\w\.]+):(\d+)\)"
8. "^(\w+)://([\w\-\.]+):?(\d+)?"

## Regex Crossword

### Beginner

1. HELP
2. BOBE
3. OOOO
4. **//
5. 1984

### Intermediate

1. ATOWEL
2. WALKER
3. FORTY-TWO
4. DONTPANIC
5. TURNOFFANDON

### Experienced

1. FOODFITFORTAKING
2. REGULAREXPRESSIONOFTHEDAY
3. ANSWERTHEGREATQUESTIONNOW
4. ?!!!?!?!?!!!?!!!?!?!?!!!?
5. TOBEORNOTTOBEAREGULARREGX

### Palindromeda

1. NETEWETEN
2. GELEYELEG
3. RATSABUTTUBASTAR
4. TRAPRAJAAJARPART
5. STEPTIMEEMITPETS

### Double Cross

1. ET
2. A-TEAM
3. NOSFERATU
4. PULP FICTION
5. FROM RUSSIA WITH LOVE

### Cites

1. LIMA
2. LONDINIUM
3. NEW YORK
4. KUALA LUMPUR
5.  COPENHAGEN 

## 14.3.5 Grouping and backreferences

```{r}

str_view(fruit, "(..)\\1", match = TRUE)
```

### 14.3.5.1 Exercises

1. Describe, in words, what these expressions will match:
    1. `(.)\1\1`  
    Any character this appears 3 times in a row "aaa"
    2. `"(.)(.)\\2\\1"`  
    Four character palindromes "anna"
    3. `(..)\1`  
    2 characters that are repeated once "abab"
    4. `"(.).\\1.\\1"`  
    Any character, any character, the first character, any character, and the first character "ataga"
    5. `"(.)(.)(.).*\\3\\2\\1"`  
    Any 3 characters, any remaining characters, the first 3 characters backwards "abcttttcba"
    
2. Construct regular expressions to match words that:
    1. Start and end with the same character.  
    `"^(.).*\\1$"`
    2. Contain a repeated pair of letters (e.g. “church” contains “ch” repeated twice.)  
    `"(..).*\\1"`
    3. Contain one letter repeated in at least three places (e.g. “eleven” contains three “e”s.)  
    `"(.).*\\1.*\\1"`