---
title: "09_27_2017"
author: "John D."
date: "September 26, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

# 23 Model basics

### 23.1.1 Prerequisites

```{r}
library(tidyverse)

library(modelr)
options(na.action = na.warn)
sim1_mod <- lm(y ~ x, data = sim1)
```

## 23.3 Visualising models

### 23.3.1 Predictions

```{r}
grid <- sim1 %>% 
  data_grid(x) 
grid

sim1_mod <- lm(y ~ x, data = sim1)

grid <- grid %>% 
  add_predictions(sim1_mod) 
grid

ggplot(sim1, aes(x)) +
  geom_point(aes(y = y)) +
  geom_line(aes(y = pred), data = grid, colour = "blue", size = 1)

```

### 23.3.2 Residuals

```{r}
sim1 <- sim1 %>% 
  add_residuals(sim1_mod)
sim1

ggplot(sim1, aes(resid)) + 
  geom_freqpoly(binwidth = 0.5)

ggplot(sim1, aes(x, resid)) + 
  geom_ref_line(h = 0) +
  geom_point() 
```

### 23.3.3 Exercises

1. Instead of using `lm()` to fit a straight line, you can use `loess()` to fit a smooth curve. Repeat the process of model fitting, grid generation, predictions, and visualisation on `sim1` using `loess()` instead of `lm()`. How does the result compare to `geom_smooth()`?

```{r}
rm(sim1)

grid <- sim1 %>% 
  data_grid(x) 
grid

sim1_mod_a <- lm(y ~ x, data = sim1)
sim1_mod_b <- loess(y ~ x, data = sim1)


grid <- grid %>% 
  add_predictions(sim1_mod_a, var = "lm")
grid <- grid %>% 
  add_predictions(sim1_mod_b, var = "loess")
grid

# plot

ggplot(sim1, aes(x)) +
  geom_point(aes(y = y)) +
  geom_line(aes(y = lm), data = grid, colour = "red", size = 1) +
  geom_line(aes(y = loess), data = grid, colour = "blue", size = 1)

ggplot(sim1, aes(x)) +
  geom_point(aes(y = y)) +
  geom_line(aes(y = lm), data = grid, colour = "red", size = 1) +
  geom_line(aes(y = loess), data = grid, colour = "darkblue", size = 3) +
  geom_line(aes(y = y), colour = "orange", size = 1) +
  geom_smooth(aes(y = y), colour = "lightblue", size = 1)

# redo with functions listed in problem 2

grid <- sim1 %>% 
  data_grid(x) 
grid

sim1_lm <- lm(y ~ x, data = sim1)
sim1_loess <- loess(y ~ x, data = sim1)

grid <- grid %>% gather_predictions(sim1_lm,sim1_loess)
grid

ggplot(sim1, aes(x)) +
  geom_point(aes(y = y)) +
  geom_line(aes(y = pred, color = model), data = grid, size = 1)

```

`geom_smooth()` fits a line using the model method 'loess'. They overlap

2. `add_predictions()` is paired with `gather_predictions()` and `spread_predictions()`. How do these three functions differ?

All work on a data frame. `add_prediction()` adds a single new column, default name 'pred', to the input data frame. `spread_predictions()` adds one column for each model. `gather_prections()` adds two columns .model and .pred, and repeats the input rows for each model. `gather_prections()` very useful for visualizing multiple lines on the same plot with labels.

```{r}
df <- tibble::data_frame(
  x = sort(runif(100)),
  y = 5 * x + 0.5 * x ^ 2 + 3 + rnorm(length(x))
)
plot(df)

m1 <- lm(y ~ x, data = df)
grid <- data.frame(x = seq(0, 1, length = 10))
grid %>% add_predictions(m1)

m2 <- lm(y ~ poly(x, 2), data = df)
grid %>% spread_predictions(m1, m2)
grid %>% gather_predictions(m1, m2)
```

3. What does `geom_ref_line()` do? What package does it come from? Why is displaying a reference line in plots showing residuals useful and important?

Add a reference line. From the modelr package, but modifies ggplot objects. Displaying a reference line in plots showing residuals is usefuly because it gives you a reference for how far from 0 the residuals are.

4. Why might you want to look at a frequency polygon of absolute residuals? What are the pros and cons compared to looking at the raw residuals?

Gives you a better overall view of how the model is performing. By doing so you lose information on if the model is over or under predicting.

## 23.4 Formulas and model families

```{r}
df <- tribble(
  ~y, ~x1, ~x2,
  4, 2, 5,
  5, 1, 6
)

model_matrix(df, y ~ x1)

model_matrix(df, y ~ x1 - 1)

model_matrix(df, y ~ x1 + x2)

```

### 23.4.1 Categorical variables

```{r}
df <- tribble(
  ~ sex, ~ response,
  "male", 1,
  "female", 2,
  "male", 1
)

model_matrix(df, response ~ sex)

ggplot(sim2) + 
  geom_point(aes(x, y))

mod2 <- lm(y ~ x, data = sim2)

grid <- sim2 %>% 
  data_grid(x) %>% 
  add_predictions(mod2)
grid

ggplot(sim2, aes(x)) + 
  geom_point(aes(y = y)) +
  geom_point(data = grid, aes(y = pred), colour = "red", size = 4)

tibble(x = "e") %>% 
  add_predictions(mod2)
```

### 23.4.2 Interactions (continuous and categorical)

```{r}
sim3
ggplot(sim3, aes(x1, y)) + 
  geom_point(aes(colour = x2))

mod1 <- lm(y ~ x1 + x2, data = sim3)
mod2 <- lm(y ~ x1 * x2, data = sim3)

grid <- sim3 %>% 
  data_grid(x1, x2) %>% 
  gather_predictions(mod1, mod2)
grid

ggplot(sim3, aes(x1, y, colour = x2)) + 
  geom_point() + 
  geom_line(data = grid, aes(y = pred)) + 
  facet_wrap(~ model)

sim3 <- sim3 %>% 
  gather_residuals(mod1, mod2)

ggplot(sim3, aes(x1, resid, colour = x2)) + 
  geom_point() + 
  facet_grid(model ~ x2)
```

### 23.4.3 Interactions (two continuous)

```{r}
mod1 <- lm(y ~ x1 + x2, data = sim4)
mod2 <- lm(y ~ x1 * x2, data = sim4)

grid <- sim4 %>% 
  data_grid(
    x1 = seq_range(x1, 5), 
    x2 = seq_range(x2, 5) 
  ) %>% 
  gather_predictions(mod1, mod2)
grid

seq_range(c(0.0123, 0.923423), n = 5)

seq_range(c(0.0123, 0.923423), n = 5, pretty = TRUE)

x1 <- rcauchy(100)
seq_range(x1, n = 5)

seq_range(x1, n = 5, trim = 0.10)

seq_range(x1, n = 5, trim = 0.25)

seq_range(x1, n = 5, trim = 0.50)

x2 <- c(0, 1)
seq_range(x2, n = 5)

seq_range(x2, n = 5, expand = 0.10)

seq_range(x2, n = 5, expand = 0.25)

seq_range(x2, n = 5, expand = 0.50)

ggplot(grid, aes(x1, x2)) + 
  geom_tile(aes(fill = pred)) + 
  facet_wrap(~ model)

ggplot(grid, aes(x1, pred, colour = x2, group = x2)) + 
  geom_line() +
  facet_wrap(~ model)

ggplot(grid, aes(x2, pred, colour = x1, group = x1)) + 
  geom_line() +
  facet_wrap(~ model)
```

### 23.4.4 Transformations

```{r}
df <- tribble(
  ~y, ~x,
   1,  1,
   2,  2, 
   3,  3
)
model_matrix(df, y ~ x^2 + x)

model_matrix(df, y ~ I(x^2) + x)

model_matrix(df, y ~ poly(x, 2))

library(splines)
model_matrix(df, y ~ ns(x, 2))

sim5 <- tibble(
  x = seq(0, 3.5 * pi, length = 50),
  y = 4 * sin(x) + rnorm(length(x))
)
sim5

ggplot(sim5, aes(x, y)) +
  geom_point()

mod1 <- lm(y ~ ns(x, 1), data = sim5)
mod2 <- lm(y ~ ns(x, 2), data = sim5)
mod3 <- lm(y ~ ns(x, 3), data = sim5)
mod4 <- lm(y ~ ns(x, 4), data = sim5)
mod5 <- lm(y ~ ns(x, 5), data = sim5)

grid <- sim5 %>% 
  data_grid(x = seq_range(x, n = 50, expand = 0.1)) %>% 
  gather_predictions(mod1, mod2, mod3, mod4, mod5, .pred = "y")
grid

ggplot(sim5, aes(x, y)) + 
  geom_point() +
  geom_line(data = grid, colour = "red") +
  facet_wrap(~ model)
```

### 23.4.5 Exercises

1. What happens if you repeat the analysis of `sim2` using a model without an intercept. What happens to the model equation? What happens to the predictions?
```{r}

model_matrix(df, y ~ x)
model_matrix(df, y ~ x - 1)

ggplot(sim2) + 
  geom_point(aes(x, y))
mod1 <- lm(y ~ x, data = sim2)
mod2 <- lm(y ~ x - 1, data = sim2)

mod1
mod2
grid <- sim2 %>% 
  data_grid(x) %>% 
  gather_predictions(mod1,mod2)
grid

ggplot(sim2, aes(x)) + 
  geom_point(aes(y = y)) +
  geom_point(data = grid, aes(y = pred), colour = "red", size = 4) +
  facet_wrap(~model)
```

Nothing really changed. Intercept did not make a difference, at least in my case

2. Use `model_matrix()` to explore the equations generated for the models I fit to `sim3` and `sim4`. Why is `*` a good shorthand for interaction?

```{r}
model_matrix(y ~ x1 + x2, data = sim3)
model_matrix(y ~ x1 * x2, data = sim3)
model_matrix(y ~ x1 + x2, data = sim4)
model_matrix(y ~ x1 * x2, data = sim4)
```

It would be a pain to write out the whole equation explicitly

3. Using the basic principles, convert the formulas in the following two models into functions. (Hint: start by converting the categorical variable into 0-1 variables.)

```{r}
mod1 <- lm(y ~ x1 + x2, data = sim3)
mod2 <- lm(y ~ x1 * x2, data = sim3)
mod1
mod2

mod1 = a_0 + a_1 * is_a + a_2 * is_b + a_3 * is_c + a_4 * is_d
mod2 = a_0 + a_1 * x1 + a_2 * x2 + a_12 * x1 * x2
# Unsure about mod 2. I feel like it's going to be long.
```

4. For `sim4`, which of `mod1` and `mod2` is better? I think `mod2` does a slightly better job at removing patterns, but it’s pretty subtle. Can you come up with a plot to support my claim?

```{r}
sim4
mod1 <- lm(y ~ x1 + x2, data = sim4)
mod2 <- lm(y ~ x1 * x2, data = sim4)
sim4 <- sim4 %>% 
  gather_residuals(mod1, mod2)
sim4

ggplot(sim4, aes(x1, resid)) + 
  geom_point() + 
  facet_grid(model ~ x2)

ggplot(sim4, aes(x1, resid)) + 
  geom_point() + 
  facet_grid(~model)

ggplot(sim4, aes(resid)) + 
  geom_freqpoly(binwidth = 0.5) +
  facet_grid(~model)
```
mod2 has a nicer looking freqpoly. More centered. I'm not good at statistics.

## 23.5 Missing values

```{r}
df <- tribble(
  ~x, ~y,
  1, 2.2,
  2, NA,
  3, 3.5,
  4, 8.3,
  NA, 10
)

mod <- lm(y ~ x, data = df)

mod <- lm(y ~ x, data = df, na.action = na.exclude)
nobs(mod)
```

## 23.6 Other model families