---
title: "09_06_2017"
author: "John D."
date: "September 13, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = T)
```

21.1-21.4

# 21 Iteration

## 21.1 Introduction

### 21.1.1 Prerequisites

```{r}
library(tidyverse)
```

## 21.2 For loops

```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
df

median(df$a)
median(df$b)
median(df$c)
median(df$d)

output <- vector("double", ncol(df))  # 1. output
for (i in seq_along(df)) {            # 2. sequence
  output[[i]] <- median(df[[i]])      # 3. body
}
output

y <- vector("double", 0)
seq_along(y)

1:length(y)
```

### 21.2.1 Exercises

1. Write for loops to:
    1. Compute the mean of every column in `mtcars`.
    2. Determine the type of each column in `nycflights13::flights`.
    3. Compute the number of unique values in each column of `iris`.
    4. Generate 10 random normals for each of μ= −10, 0, 10, and 100

```{r}
# 1

output <- vector("double", ncol(mtcars))
for (i in seq_along(mtcars)) {
  output[[i]] <- mean(mtcars[[i]])
}
output

# 2
output <- vector("character", ncol(nycflights13::flights))
for (i in seq_along(nycflights13::flights)) {
  output[[i]] <- typeof(nycflights13::flights[[i]])
}
output

# 3

output <- vector("integer", ncol(iris))
for (i in seq_along(iris)) {
  output[[i]] <- length(unique(iris[[i]]))
}
output

# 4

means <- c(-10,0,10,100)
output <- matrix(NA, nrow = 10, ncol = 4)

for (i in seq_along(means)) {
  output[,i] <- rnorm(10, mean = means[i])
}
output
```

2. Eliminate the for loop in each of the following examples by taking advantage of an existing function that works with vectors:

```{r}
# BEFORE
out <- ""
for (x in letters) {
  out <- stringr::str_c(out, x)
}
out

# AFTER
out <- stringr::str_c(letters, collapse = "")
out

# BEFORE
x <- sample(100)
sd <- 0
for (i in seq_along(x)) {
  sd <- sd + (x[i] - mean(x)) ^ 2
}
sd <- sqrt(sd / (length(x) - 1))
sd

# AFTER
sd <- sd(x)
sd

# BEFORE
x <- runif(100)
out <- vector("numeric", length(x))
out[1] <- x[1]
for (i in 2:length(x)) {
  out[i] <- out[i - 1] + x[i]
}
out

#AFTER
cumsum(x)

```

3. Combine your function writing and for loop skills:
    1. Write a for loop that `prints()` the lyrics to the children’s song “Alice the camel”.
    2. Convert the nursery rhyme “ten in the bed” to a function. Generalise it to any number of people in any sleeping structure.
    3. Convert the song “99 bottles of beer on the wall” to a function. Generalise to any number of any vessel containing any liquid on any surface.
    
```{r}
# I'm just gonna do number 3

beer_song <- function(n = 99, vessel = "bottle", liquid = "beer", surface = "wall"){
  for(i in n:1){
    cat(i, " ", vessel, "(s) of ", liquid, " on the ", surface, ".\n",
        i, " ", vessel, "(s) of ", liquid, ".\n",
        "You take one down, pass it around,\n", 
        i-1, " ", vessel, "s of ", liquid, " on the ", surface, ".\n\n",
        sep = "")
  }
  cat(0, " ", vessel, "s of ", liquid, " on the ", surface, ".\n", 
      0, " ", vessel, "(s) of ", liquid, ".\n",
      "The song is over, go home. \n",
      sep = "")
}
beer_song(10, "handle", "vodka", "table")
```

4. It’s common to see for loops that don’t preallocate the output and instead increase the length of a vector at each step: How does this affect performance? Design and execute an experiment.

```{r}
x <- sapply(100000, rnorm)
system.time({
  output <- vector("integer", 0)
for (i in seq_along(x)) {
  output <- c(output, lengths(x[[i]]))
}
output
})

system.time({
  output2 <- vector("integer", length(x))
  for (i in seq_along(output2)) {
  output2[[i]] <- lengths(x[[i]])
}
})

all.equal(output,output2)
```

## 21.3 For loop variations

### 21.3.1 Modifying an existing object

```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}

df$a <- rescale01(df$a)
df$b <- rescale01(df$b)
df$c <- rescale01(df$c)
df$d <- rescale01(df$d)

df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

for (i in seq_along(df)) {
  df[[i]] <- rescale01(df[[i]])
}

```

### 21.3.2 Looping patterns

```{r}
results <- vector("list", length(x))
names(results) <- names(x)

for (i in seq_along(x)) {
  name <- names(x)[[i]]
  value <- x[[i]]
}
```

### 21.3.3 Unknown output length

```{r}
means <- c(0, 1, 2)

output <- double()
for (i in seq_along(means)) {
  n <- sample(100, 1)
  output <- c(output, rnorm(n, means[[i]]))
}
str(output)

out <- vector("list", length(means))
for (i in seq_along(means)) {
  n <- sample(100, 1)
  out[[i]] <- rnorm(n, means[[i]])
}
str(out)
str(unlist(out))

```

### 21.3.4 Unknown sequence length

```{r}
while (condition) {
  # body
}

for (i in seq_along(x)) {
  # body
}

i <- 1
while (i <= length(x)) {
  # body
  i <- i + 1 
}

flip <- function() sample(c("T", "H"), 1)

flips <- 0
nheads <- 0

while (nheads < 3) {
  if (flip() == "H") {
    nheads <- nheads + 1
  } else {
    nheads <- 0
  }
  flips <- flips + 1
}
flips


```

### 21.3.5 Exercises

1. Imagine you have a directory full of CSV files that you want to read in. You have their paths in a vector, `files <- dir("data/", pattern = "\\.csv$", full.names = TRUE)`, and now want to read each one with `read_csv()`. Write the for loop that will load them into a single data frame.

```{r}
# Create example files
write.csv(rnorm(100), file = "data/file_1.csv")
write.csv(rnorm(100), file = "data/file_2.csv")

# Gather file names
files <- dir("data/", pattern = "\\.csv$", full.names = TRUE)
head(files)

# Read in files
dfs <- vector("list", length(files))
for (i in seq_along(dfs)){
  dfs[[i]] <- read.csv(files[[i]])
}
dfs
master_df <- dplyr::bind_rows(dfs)
master_df
```

2. What happens if you use `for (nm in names(x))` and `x` has no names? What if only some of the elements are named? What if the names are not unique?

```{r}
x <- 1:5
names(x)

for(nm in names(x)){
  print(x[[nm]])
}

names(x)[1:3] <- c("a","b","c")
names(x)

for(nm in names(x)){
  print(x[[nm]])
}

names(x)[1:5] <- c("a","b","c","a","e")
names(x)

for(nm in names(x)){
  print(x[[nm]])
}
# returns first element with that name
```

3. Write a function that prints the mean of each numeric column in a data frame, along with its name. For example, `show_mean(iris)` would print:

```{r}
show_mean <- function(df){
  for(i in seq_along(df)){
    if (is.numeric(df[[i]])){
      cat(names(df[i]),": ", round(mean(df[[i]]), 2), "\n", sep = "")
    }
  }
}
show_mean(iris)
```

4. What does this code do? How does it work?

```{r}
trans <- list( 
  disp = function(x) x * 0.0163871,
  am = function(x) {
    factor(x, labels = c("auto", "manual"))
  }
)
for (var in names(trans)) {
  mtcars[[var]] <- trans[[var]](mtcars[[var]])
}
names(trans)
mtcars
```

Allows us to use names and apply actions to specific columns

## 21.4 For loops vs. functionals

```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

output <- vector("double", length(df))
for (i in seq_along(df)) {
  output[[i]] <- mean(df[[i]])
}
output

col_mean <- function(df) {
  output <- vector("double", length(df))
  for (i in seq_along(df)) {
    output[i] <- mean(df[[i]])
  }
  output
}

col_median <- function(df) {
  output <- vector("double", length(df))
  for (i in seq_along(df)) {
    output[i] <- median(df[[i]])
  }
  output
}
col_sd <- function(df) {
  output <- vector("double", length(df))
  for (i in seq_along(df)) {
    output[i] <- sd(df[[i]])
  }
  output
}

f1 <- function(x) abs(x - mean(x)) ^ 1
f2 <- function(x) abs(x - mean(x)) ^ 2
f3 <- function(x) abs(x - mean(x)) ^ 3

f <- function(x, i) abs(x - mean(x)) ^ i


col_summary <- function(df, fun) {
  out <- vector("double", length(df))
  for (i in seq_along(df)) {
    out[i] <- fun(df[[i]])
  }
  out
}

col_summary(df, median)
col_summary(df, mean)
```

### 21.4.1 Exercises

1. Read the documentation for `apply()`. In the 2d case, what two for loops does it generalise?

Performs a for loop for either every row or every column

2. Adapt `col_summary()` so that it only applies to numeric columns You might want to start with an `is_numeric()` function that returns a logical vector that has a TRUE corresponding to each numeric column.

```{r}
col_summary <- function(df, fun) {
  df <- select_if(df, is.numeric)
  out <- vector("double", length(df))
  for (i in seq_along(df)) {
    out[i] <- fun(df[[i]])
  }
  names(out) <- names(df)
  out
}
col_summary(iris, mean)
col_summary(iris, median)
```

