---
title: "09_13_2017"
author: "John D."
date: "September 13, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = T)
```

# 21 Iteration

```{r}
library(tidyverse)
```

## 21.5 The map functions

```{r}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

map_dbl(df, mean)
map_dbl(df, median)
map_dbl(df, sd)

df %>% map_dbl(mean)
df %>% map_dbl(median)
df %>% map_dbl(sd)

map_dbl(df, mean, trim = 0.5)

z <- list(x = 1:3, y = 4:5)
map_int(z, length)
```

### 21.5.1 Shortcuts

```{r}
models <- mtcars %>% 
  split(.$cyl) %>% 
  map(function(df) lm(mpg ~ wt, data = df))

models <- mtcars %>% 
  split(.$cyl) %>% 
  map(~lm(mpg ~ wt, data = .))

models <- list(mtcars) %>% 
  #split(.$cyl) %>% 
  map(~lm(mpg ~ wt, data = .))

models %>% 
  map(summary) %>% 
  map_dbl(~.$r.squared)

models %>% 
  map(summary) %>% 
  map_dbl("r.squared")

x <- list(list(1, 2, 3), list(4, 5, 6), list(7, 8, 9))
x %>% map_dbl(2)
```

### 21.5.2 Base R

```{r}
x1 <- list(
  c(0.27, 0.37, 0.57, 0.91, 0.20),
  c(0.90, 0.94, 0.66, 0.63, 0.06), 
  c(0.21, 0.18, 0.69, 0.38, 0.77)
)
x2 <- list(
  c(0.50, 0.72, 0.99, 0.38, 0.78), 
  c(0.93, 0.21, 0.65, 0.13, 0.27), 
  c(0.39, 0.01, 0.38, 0.87, 0.34)
)

threshold <- function(x, cutoff = 0.8) x[x > cutoff]

x1 %>% sapply(threshold) %>% str()

x2 %>% sapply(threshold) %>% str()

map(x2, threshold) %>% str()

### in x1, first entry has 1 that passes, second has 2, and 3rd has none.
### in x2, each entry has 1 value that passes
```

### 21.5.3 Exercises

1. Write code that uses one of the map functions to:
    1. Compute the mean of every column in `mtcars`.
    2. Determine the type of each column in `nycflights13::flights`.
    3. Compute the number of unique values in each column of `iris`.
    4. Generate 10 random normals for each of μ = -10, 0, 10, and 100

```{r}
# 1
mtcars %>% map_dbl(mean)

# 2
nycflights13::flights %>% map_chr(typeof)

# 3
iris %>% map_int(~length(unique(.)))

# 4
means <- c(-10, 0, 10, 100)
map(means, rnorm, n = 10)
  
# Works because second argument to rnrom is mean and n has already been used
```

2. How can you create a single vector that for each column in a data frame indicates whether or not it’s a factor?

```{r}
map_lgl(iris, is.factor)
```

3. What happens when you use the map functions on vectors that aren’t lists? What does `map(1:5, runif)` do? Why?

```{r}
map(1:5, runif)

```

Applies function to each iteration and creates a list ie(runif(1), runif(2)).

4. What does `map(-2:2, rnorm, n = 5)` do? Why? What does `map_dbl(-2:2, rnorm, n = 5)` do? Why?

```{r}
map(-2:2, rnorm, n = 5)
```

Creates a list of lists. For each iteration, the iteration value is the mean and 5 values are computed with `rnorm`

```{r}
map_dbl(-2:2, rnorm, n = 5)
```

Causes an error because the output is not a vector of length 1

5. Rewrite `map(x, function(df) lm(mpg ~ wt, data = df))` to eliminate the anonymous function.

```{r}
x <- mtcars
# BEFORE
map(x, function(df) lm(mpg ~ wt, data = df))

#AFTER
map(list(mtcars), ~ lm(mpg ~ wt, data = .))

# map takes lists as input so convert x to list to start
# then use '.' as a pronoun to refer to current list item
# perform the model function on current list item
```

## 21.6 Dealing with failure

```{r}
safe_log <- safely(log)
str(safe_log(10))
str(safe_log("a"))

x <- list(1, 10, "a")
y <- x %>% map(safely(log))
str(y)

y <- y %>% transpose()
str(y)

is_ok <- y$error %>% map_lgl(is_null)
x[!is_ok]

y$result[is_ok] %>% flatten_dbl()

x <- list(1, 10, "a")
x %>% map_dbl(possibly(log, NA_real_))

x <- list(1, -1)
x %>% map(quietly(log)) %>% str()
```

## 21.7 Mapping over multiple arguments

```{r}
mu <- list(5, 10, -3)
mu %>% 
  map(rnorm, n = 5) %>% 
  str()

sigma <- list(1, 5, 10)
seq_along(mu) %>% 
  map(~rnorm(5, mu[[.]], sigma[[.]])) %>% 
  str()

map2(mu, sigma, rnorm, n = 5) %>% str()


n <- list(1, 3, 5)
args1 <- list(n, mu, sigma)
args1 %>%
  pmap(rnorm) %>% 
  str()

args2 <- list(mean = mu, sd = sigma, n = n)
args2 %>% 
  pmap(rnorm) %>% 
  str()

params <- tribble(
  ~mean, ~sd, ~n,
    5,     1,  1,
   10,     5,  3,
   -3,    10,  5
)

params %>% 
  pmap(rnorm)
```

### 21.7.1 Invoking different functions

```{r}
f <- c("runif", "rnorm", "rpois")
param <- list(
  list(min = -1, max = 1), 
  list(sd = 5), 
  list(lambda = 10)
)
str(param)

invoke_map(f, param, n = 5) %>% str()

sim <- tribble(
  ~f,      ~params,
  "runif", list(min = -1, max = 1),
  "rnorm", list(sd = 5),
  "rpois", list(lambda = 10)
)
a <- sim %>% 
  mutate(sim = invoke_map(f, params, n = 10))

a$sim
```

## 21.8 Walk

```{r}
x <- list(1, "a", 3)

x %>% 
  walk(print)

library(ggplot2)
plots <- mtcars %>% 
  split(.$cyl) %>% 
  map(~ggplot(., aes(mpg, wt)) + geom_point())
paths <- stringr::str_c(names(plots), ".pdf")

pwalk(list(paths, plots), ggsave, path = "plots")
```

## 21.9 Other patterns of for loops

### 21.9.1 Predicate functions

```{r}
iris %>% 
  keep(is.factor) %>% 
  str()

iris %>% 
  discard(is.factor) %>% 
  str()

x <- list(1:5, letters, list(10))

x %>% 
  some(is_character)

x %>% 
  every(is_vector)

x %>% 
  every(is_character)

x <- sample(10)
x

x %>% 
  detect(~ . > 5)

x %>% 
  detect_index(~ . > 5)

x %>% 
  head_while(~ . > 5)

x %>% 
  tail_while(~ . > 5)
```

### 21.9.2 Reduce and accumulate

```{r}
dfs <- list(
  age = tibble(name = "John", age = 30),
  sex = tibble(name = c("John", "Mary"), sex = c("M", "F")),
  trt = tibble(name = "Mary", treatment = "A")
)

dfs %>% reduce(full_join)

vs <- list(
  c(1, 3, 5, 6, 10),
  c(1, 2, 3, 7, 8, 10),
  c(1, 2, 3, 4, 8, 9, 10)
)

vs %>% reduce(intersect)

x <- sample(10)
x

x %>% accumulate(`+`)


```

### 21.9.3 Exercises

1. Implement your own version of `every()` using a for loop. Compare it with `purrr::every()`. What does purrr’s version do that your version doesn’t?

```{r}
x <- list(1:5, letters, list(10))
x %>% 
  every(is_vector)
x %>% 
  every(is_character)

every2 <- function(x, fun, ...){
  for (i in seq_along(x)){
    if(!fun(x, ...)) return(FALSE)
  }
  return(TRUE)
}
every2(x, is_vector)
every2(x, is_character)
```

`every()` has extra abilities. For a function it can take a logical vector of the same length as x. Alternatively, if the elements of x are themselves lists of objects, a string indicating the name of a logical element in the inner lists. Only those elements where p evaluates to TRUE will be modified.

2. Create an enhanced `col_summary()` that applies a summary function to every numeric column in a data frame.

```{r}

col_summary <- function(df, fun) {
  out <- vector("double", length(df))
  for (i in seq_along(df)) {
    out[i] <- fun(df[[i]])
  }
  out
}

col_summary2 <- function(df, fun, ...) {
  map(keep(df, is.numeric), fun, ...)
}

col_summary(iris, median)
col_summary2(iris, median)
col_summary2(iris, mean)
col_summary2(iris, mean, trim = .5)

```
Is this a repeat of a question from last week? Guess want us to rewrite with new learned tools

3. A possible base R equivalent of `col_summary()` is:

```{r}
col_summary3 <- function(df, f) {
  is_num <- sapply(df[0], is.numeric)
  df_num <- df[, is_num]

  sapply(df_num, f)
}

col_summary4 <- function(df, f, ...) {
  df_num <- keep(df, is.numeric)
  map_dbl(df_num, f, ...)
}
```

But it has a number of bugs as illustrated with the following inputs:

```{r}
df <- tibble(
  x = 1:3, 
  y = 3:1,
  z = c("a", "b", "c")
)
# OK
col_summary3(df, mean)
# Has problems: don't always return numeric vector
col_summary3(df[1:2], mean)
col_summary3(df[1], mean)
col_summary3(df[0], mean)

df2 <- df[0]
df2
is_num <- sapply(df2, is.numeric)
is_num
df_num <- df2[, is_num]
```

Since the first `sapply()` had an empty tibble, it returned an empty list by default. The subsetting then fails because you are attempting to subset using a list which is an unsupported index type. The final `sapply()` never runs because the function crashes at the logical subsetting step.