---
title: "04_25_2017"
author: "John D."
date: "April 25, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r prereq}
#install.packages("tidyverse")
library(tidyverse)
```

```{r 3.2.1}
#mpg
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

Template  
`ggplot(data = <DATA>) + `
`  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))`

### 3.2.4 Exercises  

1. Run `ggplot(data = mpg)` what do you see?
```{r 3.2.4.1}
ggplot(data = mpg)
```

A blank plot is produced since no layers have been applied.

2. How many rows are in `mtcars`? How many columns?

```{r 3.2.4.2}
dim(mtcars)
```

32 rows and 11 columns

3. What does the `drv` variable describe? Read the help for`?mpg` to find out.

`?mpg` drv describes the drive type of the car. ie front-wheel

4. Make a scatterplot of `hwy` vs `cyl`

```{r 3.2.4.4}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl))
```

5. What happens if you make a scatterplot of `class` vs `drv`. Why is the plot not useful?

```{r 3.2.4.5}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv))
```

There is not a relationship between class and drive type. I personally can make use of this plot because it tells me the available drive types for each class of car.

### 3.3.1 Exercises  

1. What’s gone wrong with this code? Why are the points not blue?  

color = 'blue' needs to be outside of the aes function

2. Which variables in `mpg` are categorical? Which variables are continuous? (Hint: type `?mpg` to read the documentation for the dataset). How can you see this information when you run `mpg`?

```{r 3.3.1.2}
mpg
```

Columns of type `chr` are categorical, type `int` are continous, and type `dbl` can be either depending on the contents

3. Map a continuous variable to `color`, `size`, and `shape`. How do these aesthetics behave differently for categorical vs. continuous variables?

```{r 3.3.1.3}
# ggplot(data = mpg) +
#   geom_point(mapping = aes(x = cyl, y = hwy, shape= cty, color = year, size = cyl))
```

Shape cannot be mapped to a continous variable

4. What happens if you map the same variable to multiple aesthetics?

```{r 3.3.1.4}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = cyl, y = hwy, color = hwy, size = hwy))
```

It will apply each aesthetic appropriately according to the values in the variable

5. What does the `stroke` aesthetic do? What shapes does it work with? (Hint: use `?geom_point`)

```{r 3.3.1.5}
ggplot(mtcars, aes(wt, mpg)) +
  geom_point(shape = 21, colour = "black", fill = "white", stroke = 5)
```

Stroke changes the size of the border for the points on the plot

6. What happens if you map an aesthetic to something other than a variable name, like `aes(colour = displ < 5)`?

```{r 3.3.1.6}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl, color = displ < 5))
```

Turns the color scheme into a boolean of either true or false