---
title: "Ch.15"
author: "John D."
date: "July 20, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(forcats)
```

# Factors

## 15.2 Creating factors

```{r}
x1 <- c("Dec", "Apr", "Jan", "Mar")

x2 <- c("Dec", "Apr", "Jam", "Mar")

sort(x1)

month_levels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun", 
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)

y1 <- factor(x1, levels = month_levels)
y1
sort(y1)

y2 <- factor(x2, levels = month_levels)
y2

y2 <- parse_factor(x2, levels = month_levels)

factor(x1)

f1 <- factor(x1, levels = unique(x1))
f1

f2 <- x1 %>% factor() %>% fct_inorder()
f2

levels(f2)
```

## General Social Survey

```{r}
gss_cat

gss_cat %>%
  count(race)

ggplot(gss_cat, aes(race)) +
  geom_bar()

ggplot(gss_cat, aes(race)) +
  geom_bar() +
  scale_x_discrete(drop = FALSE)
```

### 15.3.1 Exercise

1. Explore the distribution of `rincome` (reported income). What makes the default bar chart hard to understand? How could you improve the plot?

```{r}
ggplot(gss_cat, aes(rincome)) +
  geom_bar() +
  scale_x_discrete(drop = FALSE) +
  theme(axis.text.x = element_text(angle = 90))

ggplot(gss_cat, aes(rincome)) +
  geom_bar() +
  scale_x_discrete(limits = rev(levels(gss_cat$rincome)), drop = FALSE) +
  theme(axis.text.x = element_text(angle = 90))
```
The money amounts go from most to least which is reverse of how most would present it. Also would be nice to have categories which do not have a dollar range grouped together at the end.

2. What is the most common `relig` in this survey? What’s the most common `partyid`?

```{r}
gss_cat %>%
  count(relig, sort = T) %>%
  head(1)

gss_cat %>%
  count(partyid, sort = T) %>%
  head(1)
```

3. Which `relig` does `denom` (denomination) apply to? How can you find out with a table? How can you find out with a visualisation?

```{r}
gss_cat %>%
  group_by(relig) %>%
  count(denom, sort = T)

gss_cat %>%
  group_by(relig) %>%
  count(denom, sort = T) %>%
  mutate(comb = paste(relig,denom, sep = "_")) %>%
  ggplot(aes(reorder(comb, order(n)), weight = n)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90))

gss_cat %>%
  filter(relig == "Protestant") %>%
  count(denom) %>%
  ggplot(aes(reorder(denom, n), n)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90))
```

## Modifying factor order

```{r}
relig_summary <- gss_cat %>%
  group_by(relig) %>%
  summarise(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )

ggplot(relig_summary, aes(tvhours, relig)) + geom_point()
ggplot(relig_summary, aes(tvhours, fct_reorder(relig, tvhours))) +
  geom_point()

relig_summary %>%
  mutate(relig = fct_reorder(relig, tvhours)) %>%
  ggplot(aes(tvhours, relig)) +
    geom_point()

rincome_summary <- gss_cat %>%
  group_by(rincome) %>%
  summarise(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )

ggplot(rincome_summary, aes(age, fct_reorder(rincome, age))) + geom_point()

ggplot(rincome_summary, aes(age, fct_relevel(rincome, "Not applicable"))) +
  geom_point()

by_age <- gss_cat %>%
  filter(!is.na(age)) %>%
  group_by(age, marital) %>%
  count() %>%
  ungroup() %>%
  group_by(age) %>%
  mutate(prop = n / sum(n))

ggplot(by_age, aes(age, prop, colour = marital)) +
  geom_line(na.rm = TRUE)

ggplot(by_age, aes(age, prop, colour = fct_reorder2(marital, age, prop))) +
  geom_line() +
  labs(colour = "marital")

gss_cat %>%
  mutate(marital = marital %>% fct_infreq() %>% fct_rev()) %>%
  ggplot(aes(marital)) +
    geom_bar()
```

### 15.4.1 Exercises

1. There are some suspiciously high numbers in tvhours. Is the mean a good summary?

```{r}
summary(gss_cat$tvhours)

gss_cat %>%
  group_by(tvhours) %>%
  count() %>%
  ggplot(aes(tvhours, weight = n)) + 
  geom_bar()
```

The mean is a good summary despite the erroneous high numbers

2. For each factor in `gss_cat` identify whether the order of the levels is arbitrary or principled

```{r}
str(gss_cat)
levels(gss_cat$marital)
levels(gss_cat$race)
levels(gss_cat$rincome)
levels(gss_cat$partyid)
levels(gss_cat$relig)
levels(gss_cat$denom)
```

Marital is arbitrary, race is arbitrary, rincome is principled, partyid is principled, relig is arbitrary, and denom is arbitrary

3. Why did moving “Not applicable” to the front of the levels move it to the bottom of the plot?

Moving "Not applicable" to the front gave it the index value of 1 which puts it at the bottom of the y axis.

## 15.5 Modifying factor levels

```{r}
gss_cat %>% count(partyid)

gss_cat %>%
  mutate(partyid = fct_recode(partyid,
    "Republican, strong"    = "Strong republican",
    "Republican, weak"      = "Not str republican",
    "Independent, near rep" = "Ind,near rep",
    "Independent, near dem" = "Ind,near dem",
    "Democrat, weak"        = "Not str democrat",
    "Democrat, strong"      = "Strong democrat"
  )) %>%
  count(partyid)

gss_cat %>%
  mutate(partyid = fct_recode(partyid,
    "Republican, strong"    = "Strong republican",
    "Republican, weak"      = "Not str republican",
    "Independent, near rep" = "Ind,near rep",
    "Independent, near dem" = "Ind,near dem",
    "Democrat, weak"        = "Not str democrat",
    "Democrat, strong"      = "Strong democrat",
    "Other"                 = "No answer",
    "Other"                 = "Don't know",
    "Other"                 = "Other party"
  )) %>%
  count(partyid)

gss_cat %>%
  mutate(partyid = fct_collapse(partyid,
    other = c("No answer", "Don't know", "Other party"),
    rep = c("Strong republican", "Not str republican"),
    ind = c("Ind,near rep", "Independent", "Ind,near dem"),
    dem = c("Not str democrat", "Strong democrat")
  )) %>%
  count(partyid)

gss_cat %>%
  mutate(relig = fct_lump(relig)) %>%
  count(relig)

gss_cat %>%
  mutate(relig = fct_lump(relig, n = 10)) %>%
  count(relig, sort = TRUE) %>%
  print(n = Inf)
```

### 15.5.1 Exercises

1. How have the proportions of people identifying as Democrat, Republican, and Independent changed over time?

```{r}
parties <- gss_cat %>%
  mutate(partyid = fct_collapse(partyid,
    other = c("No answer", "Don't know", "Other party"),
    rep = c("Strong republican", "Not str republican"),
    ind = c("Ind,near rep", "Independent", "Ind,near dem"),
    dem = c("Not str democrat", "Strong democrat")
  )) %>%
  group_by(year) %>%
  count(partyid) %>%
  mutate(prop = n / sum(n))

ggplot(parties, aes(year, prop, fill = partyid)) +
  geom_bar(stat = "identity")

ggplot(parties, aes(partyid, prop, fill = partyid)) +
  geom_bar(stat = "identity") +
  facet_wrap(~year)
```

2. How could you collapse rincome into a small set of categories?

```{r}
gss_cat %>%
  mutate(rincome = fct_collapse(rincome,
    other = c("No answer", "Don't know", "Refused", "Not applicable"),
    `$4999 or less` = c("Lt $1000", "$1000 to 2999", "$3000 to 3999", "$4000 to 4999"),
    `$5000 to 9999`= c("$5000 to 5999", "$6000 to 6999", "$7000 to 7999", "$8000 to 9999")
  )) %>%
  count(rincome)
```