---
title: "Ch.16"
author: "John D."
date: "July 20, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 16 Dates and times

## 16.1 Introduction

### 16.1.1 Prerequisites

```{r}
library(tidyverse)

library(lubridate)
library(nycflights13)
```

## Creating date/times

```{r}
today()

now()
```

### 16.2.1 From strings

```{r}
ymd("2017-01-31")
mdy("January 31st, 2017")
dmy("31-Jan-2017")
ymd(20170131)

ymd_hms("2017-01-31 20:11:59")
mdy_hm("01/31/2017 08:01")
ymd(20170131, tz = "UTC")
```

### 16.2.2 From individual components

```{r}
flights %>% 
  select(year, month, day, hour, minute)

flights %>% 
  select(year, month, day, hour, minute) %>% 
  mutate(departure = make_datetime(year, month, day, hour, minute))

make_datetime_100 <- function(year, month, day, time) {
  make_datetime(year, month, day, time %/% 100, time %% 100)
}

flights_dt <- flights %>% 
  filter(!is.na(dep_time), !is.na(arr_time)) %>% 
  mutate(
    dep_time = make_datetime_100(year, month, day, dep_time),
    arr_time = make_datetime_100(year, month, day, arr_time),
    sched_dep_time = make_datetime_100(year, month, day, sched_dep_time),
    sched_arr_time = make_datetime_100(year, month, day, sched_arr_time)
  ) %>% 
  select(origin, dest, ends_with("delay"), ends_with("time"))

flights_dt %>% 
  ggplot(aes(dep_time)) + 
  geom_freqpoly(binwidth = 86400) # 86400 seconds = 1 day

flights_dt %>% 
  filter(dep_time < ymd(20130102)) %>% 
  ggplot(aes(dep_time)) + 
  geom_freqpoly(binwidth = 600) # 600 s = 10 minutes
```

### 16.2.3 From other types

```{r}
as_datetime(today())

as_date(now())

as_datetime(60 * 60 * 10)

as_date(365 * 10 + 2)
```

### 16.2.4 Exercises

1. What happens if you parse a string that contains invalid dates?
```{r}
ymd(c("2010-10-10", "bananas"))
```

Prints with a warning message and the failed to parse item becomes NA

2. What does the `tzone` argument to `today()` do? Why is it important?

```{r}
today("America/Los_Angeles")
today("Australia/Sydney")
```

Tells the function what timezone to use when calculating today's date. Due to timezones parts of the world are on different days at the same time

3. Use the appropriate lubridate function to parse each of the following dates:

```{r}
# 1
d1 <- "January 1, 2010"
mdy(d1)
# 2
d2 <- "2015-Mar-07"
ymd(d2)
# 3
d3 <- "06-Jun-2017"
dmy(d3)
# 4
d4 <- c("August 19 (2015)", "July 1 (2015)")
mdy(d4)
# 5
d5 <- "12/30/14" # Dec 30, 2014
mdy(d5)
```

## 16.3 Date-time components

### 16.3.1 Getting components

```{r}
datetime <- ymd_hms("2016-07-08 12:34:56")
year(datetime)
month(datetime)
mday(datetime)
yday(datetime)
wday(datetime)
month(datetime, label = TRUE)
wday(datetime, label = TRUE, abbr = FALSE)

flights_dt %>% 
  mutate(wday = wday(dep_time, label = TRUE)) %>% 
  ggplot(aes(x = wday)) +
    geom_bar()

flights_dt %>% 
  mutate(minute = minute(dep_time)) %>% 
  group_by(minute) %>% 
  summarise(
    avg_delay = mean(arr_delay, na.rm = TRUE),
    n = n()) %>% 
  ggplot(aes(minute, avg_delay)) +
    geom_line()

sched_dep <- flights_dt %>% 
  mutate(minute = minute(sched_dep_time)) %>% 
  group_by(minute) %>% 
  summarise(
    avg_delay = mean(arr_delay, na.rm = TRUE),
    n = n())

ggplot(sched_dep, aes(minute, avg_delay)) +
  geom_line()

ggplot(sched_dep, aes(minute, n)) +
  geom_line()
```

### 16.3.2 Rounding

```{r}
flights_dt %>% 
  count(week = floor_date(dep_time, "day")) %>% 
  ggplot(aes(week, n)) +
    geom_line()
```

### 16.3.3 Setting components

```{r}
(datetime <- ymd_hms("2016-07-08 12:34:56"))

year(datetime) <- 2020
datetime

month(datetime) <- 01
datetime

hour(datetime) <- hour(datetime) + 1
datetime

update(datetime, year = 2020, month = 2, mday = 2, hour = 2)

ymd("2015-02-01") %>% 
  update(mday = 30)

ymd("2015-02-01") %>% 
  update(hour = 400)

flights_dt %>% 
  mutate(dep_hour = update(dep_time, yday = 1)) %>% 
  ggplot(aes(dep_hour)) +
    geom_freqpoly(binwidth = 300)
```

### 16.3.4

1. How does the distribution of flight times within a day change over the course of the year?

```{r}
flights_dt %>%
  mutate(time = hour(dep_time) * 100 + minute(dep_time),
         day = as.factor(yday
                         (dep_time))) %>%
  ggplot(aes(x = time, group = day, color = day)) +
  geom_freqpoly(binwidth = 100)
```

2. Compare dep_time, sched_dep_time and dep_delay. Are they consistent? Explain your findings.

```{r}
flights_dt %>%
  mutate(dep_time2 = sched_dep_time + dep_delay * 60) %>%
  filter(dep_time2 != dep_time) %>%
  select(dep_time2, dep_time, sched_dep_time, dep_delay)
```

3. Compare air_time with the duration between the departure and arrival. Explain your findings. (Hint: consider the location of the airport.)
```{r}
flights_dt %>%
  mutate(est_air_time = minutes(arr_time - dep_time),
         act_air_time = minutes(air_time)) %>%
  filter(act_air_time != est_air_time) %>%
  select(origin,dest,act_air_time,est_air_time)
```

The different locations of the airports can be in different timezones. The times are most likely relative to the timezone of the airport they are recorded at.

4. How does the average delay time change over the course of a day? Should you use `dep_time` or `sched_dep_time`? Why?

```{r}
flights_dt %>%
  mutate(sched_dep_hour = hour(sched_dep_time)) %>%
  group_by(sched_dep_hour) %>%
  summarise(dep_delay = mean(dep_delay)) %>%
  ggplot(aes(y = dep_delay, x = sched_dep_hour)) +
  geom_point() +
  geom_smooth()
```

Delays increase throughout the day until about 7 pm

5. On what day of the week should you leave if you want to minimise the chance of a delay?
```{r}
flights_dt %>%
  mutate(sched_dep_day = wday(sched_dep_time, label = T)) %>%
  group_by(sched_dep_day) %>%
  summarise(dep_delay = mean(dep_delay, na.rm = T)) %>%
  ggplot(aes(y = dep_delay, x = sched_dep_day)) +
  geom_bar(stat = "identity")

flights_dt %>%
  mutate(sched_arr_day = wday(sched_arr_time, label = T)) %>%
  group_by(sched_arr_day) %>%
  summarise(arr_delay = mean(arr_delay, na.rm = T)) %>%
  ggplot(aes(y = arr_delay, x = sched_arr_day)) +
  geom_bar(stat = "identity")
```

The best day is saturday

6. What makes the distribution of `diamonds$carat` and `flights$sched_dep_time` similar?

```{r}
ggplot(diamonds, aes(carat)) +
  geom_density()

ggplot(flights, aes(sched_dep_time)) +
  geom_density()
```

Humans like nice numbers hence the peaks

7. Confirm my hypothesis that the early departures of flights in minutes 20-30 and 50-60 are caused by scheduled flights that leave early. Hint: create a binary variable that tells you whether or not a flight was delayed.

```{r}
flights_dt %>%
  mutate(early = dep_delay < 0,
         minute = minute(sched_dep_time) %/% 10) %>%
  group_by(minute) %>%
  summarise(early = mean(early)) %>%
  ggplot(aes(x = minute, y = early)) +
  geom_point()
```

Hypothesis doesn't seem to be confirmed based off this result

## 16.4 Time spans

### 16.4.1 Durations

```{r}
h_age <- today() - ymd(19791014)
h_age

as.duration(h_age)

dseconds(15)
dminutes(10)
dhours(c(12, 24))
ddays(0:5)
dweeks(3)
dyears(1)

2 * dyears(1)
dyears(1) + dweeks(12) + dhours(15)

tomorrow <- today() + ddays(1)
last_year <- today() - dyears(1)


one_pm <- ymd_hms("2016-03-12 13:00:00", tz = "America/New_York")
one_pm
one_pm + ddays(1)
```

### 16.4.2 Periods

```{r}
one_pm
one_pm + days(1)

seconds(15)
minutes(10)
hours(c(12, 24))
days(7)
months(1:6)
weeks(3)
years(1)
10 * (months(6) + days(1))
days(50) + hours(25) + minutes(2)
ymd("2016-01-01") + dyears(1)
ymd("2016-01-01") + years(1)
one_pm + ddays(1)
one_pm + days(1)
flights_dt %>% 
  filter(arr_time < dep_time) 

flights_dt <- flights_dt %>% 
  mutate(
    overnight = arr_time < dep_time,
    arr_time = arr_time + days(overnight * 1),
    sched_arr_time = sched_arr_time + days(overnight * 1)
  )

flights_dt %>% 
  filter(overnight, arr_time < dep_time) 

```

### 16.4.3 Intervals

```{r}
years(1) / days(1)
next_year <- today() + years(1)
(today() %--% next_year) / ddays(1)
(today() %--% next_year) %/% days(1)
```

### 16.4.5 Exercises

1. Why is there months() but no dmonths()?

The number of days in a month changes based on the month and year

2. Explain days(overnight * 1) to someone who has just started learning R. How does it work?

If a flight arrives before it lands due to timezomes if can cause problems. Overnight is a boolean which tells you if this occurred. If this did occur, the function adds 1 day to the arrival date to correct the error, TRUE = 1 and FALSE = 0

3. Create a vector of dates giving the first day of every month in 2015. Create a vector of dates giving the first day of every month in the current year.

```{r}
ymd("2015-01-01") + months(0:11)
floor_date(today(), "year") + months(0:11)
```

4. Write a function that given your birthday (as a date), returns how old you are in years.

```{r}
age <- function(x) {
  (ymd(x) %--% today()) %/% years(1)
}
age("1993-10-10")
```

5. Why can’t `(today() %--% (today() + years(1)) / months(1)` work?

```{r}
(today() %--% (today() + years(1)) / months(1))
```

You forget a parenthesis...

## 16.5 Time zones

```{r}
Sys.timezone()
length(OlsonNames())
head(OlsonNames())

(x1 <- ymd_hms("2015-06-01 12:00:00", tz = "America/New_York"))
(x2 <- ymd_hms("2015-06-01 18:00:00", tz = "Europe/Copenhagen"))
(x3 <- ymd_hms("2015-06-02 04:00:00", tz = "Pacific/Auckland"))

x1 - x2
x1 - x3

x4 <- c(x1, x2, x3)
x4

x4a <- with_tz(x4, tzone = "Australia/Lord_Howe")
x4a

x4a - x4

x4b <- force_tz(x4, tzone = "Australia/Lord_Howe")
x4b
x4b - x4

```

