---
title: "06_28_2017"
author: "John D."
date: "June 28, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(nycflights13)
library(RcppRoll)
```

## 13.3 Keys

```{r}
planes %>% 
  count(tailnum) %>% 
  filter(n > 1)

weather %>% 
  count(year, month, day, hour, origin) %>% 
  filter(n > 1)

flights %>% 
  count(year, month, day, flight) %>% 
  filter(n > 1)

flights %>% 
  count(year, month, day, tailnum) %>% 
  filter(n > 1)
```

### 13.3.1 Exercises

1. Add a surrogate key to `flights`.

```{r}
sur_flights <- flights %>%
  mutate(surrogate = row_number()) %>%
  select(surrogate,year:time_hour)
sur_flights
```

2. Identify the keys in the following datasets

```{r}
#Lahman::Batting
summary(Lahman::Batting)
Lahman::Batting %>%
  count(playerID,yearID,stint) %>%
  filter(n > 1)

#babynames::babynames
summary(babynames::babynames)
babynames::babynames %>%
  count(year,name, sex) %>%
  filter(nn > 1)

# nasaweather::atmos
summary(nasaweather::atmos)
nasaweather::atmos %>%
  count(lat,long, year, month) %>%
  filter(n > 1)

# fueleconomy::vehicles
summary(fueleconomy::vehicles)
head(fueleconomy::vehicles)
fueleconomy::vehicles %>%
  count(id) %>%
  filter(n > 1)

# ggplot2::diamonds
summary(ggplot2::diamonds)
ggplot2::diamonds %>%
  count(depth,clarity,carat,color,price,cut,table,x,y,z) %>%
  filter(n > 1)
### Impossible exact duplicates
surg_diamonds <- diamonds %>%
  mutate(surrogate_key = row_number()) %>%
  select(surrogate_key, carat:z)
surg_diamonds

## or account for duplicates
ggplot2::diamonds %>%
  count(depth,clarity,carat,color,price,cut,table,x,y,z) %>%
  filter(n > 2)
# Can't
```

3. Draw a diagram illustrating the connections between the `Batting`, `Master`, and `Salaries` tables in the Lahman package. Draw another diagram that shows the relationship between `Master`, `Managers`, `AwardsManagers`.

How would you characterise the relationship between the Batting, Pitching, and Fielding tables?

I'm gonna skip this one

## 13.4 Mutating joins

```{r}
flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)
flights2

flights2 %>%
  select(-origin, -dest) %>% 
  left_join(airlines, by = "carrier")

flights2 %>%
  select(-origin, -dest) %>% 
  mutate(name = airlines$name[match(carrier, airlines$carrier)])

```

### 13.4.1 Understanding joins
```{r}
x <- tribble(
  ~key, ~val_x,
     1, "x1",
     2, "x2",
     3, "x3"
)
y <- tribble(
  ~key, ~val_y,
     1, "y1",
     2, "y2",
     4, "y3"
)
```

### 13.4.2 Inner join

```{r}
x
y
x %>% 
  inner_join(y, by = "key")
```

### 13.4.3 Outer joins

### 13.4.4 Duplicate keys

```{r}
x <- tribble(
  ~key, ~val_x,
     1, "x1",
     2, "x2",
     2, "x3",
     1, "x4"
)
y <- tribble(
  ~key, ~val_y,
     1, "y1",
     2, "y2"
)
x
y
left_join(x, y, by = "key")

x <- tribble(
  ~key, ~val_x,
     1, "x1",
     2, "x2",
     2, "x3",
     3, "x4"
)
y <- tribble(
  ~key, ~val_y,
     1, "y1",
     2, "y2",
     2, "y3",
     3, "y4"
)

x
y
left_join(x, y, by = "key")
```

### 13.4.5 Defining the key columns

```{r}
flights2 %>% 
  left_join(weather)

flights2 %>% 
  left_join(planes, by = "tailnum")

flights2 %>% 
  left_join(airports, c("dest" = "faa"))

flights2 %>% 
  left_join(airports, c("origin" = "faa"))
```

### 13.4.6 Exercises

1. Compute the average delay by destination, then join on the `airports` data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:

(Don’t worry if you don’t understand what semi_join() does — you’ll learn about it next.)

You might want to use the size or colour of the points to display the average delay for each airport.
```{r}
flights3 <- flights %>%
  group_by(dest) %>%
  mutate(avg_delay = mean(arr_delay, na.rm = T)) %>%
  select(dest,avg_delay) %>%
  inner_join(airports, by = c(dest = "faa")) 

flights3 %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point(aes(color = avg_delay)) +
    coord_quickmap()
```

2. Add the location of the origin and destination (i.e. the `lat` and `lon`) to flights.

```{r}
flights4 <- flights2 %>%
  left_join(airports %>% select(faa,lat,lon), by = c(origin = "faa")) %>%
  left_join(airports%>% select(faa,lat,lon), by = c(dest = "faa")) %>%
  rename(origin.lat = lat.x ,  dest.lat = lat.y, origin.lon = lon.x, dest.lon = lon.y)
flights4
```

3. Is there a relationship between the age of a plane and its delays?

```{r}
flights %>%
  group_by(tailnum) %>%
  mutate(avg_delay = mean(arr_delay, na.rm = T)) %>%
  select(tailnum,avg_delay) %>%
  unique() %>%
  inner_join(planes) %>%
  mutate(age = 2013-year) %>%
  ggplot(aes(age,avg_delay)) +
    geom_point()

flights %>%
  group_by(tailnum) %>%
  mutate(avg_delay = mean(arr_delay, na.rm = T)) %>%
  select(tailnum,avg_delay) %>%
  unique() %>%
  inner_join(planes) %>%
  mutate(age = 2013-year) %>%
  ggplot(aes(age,avg_delay)) +
    geom_line()
```

4. What weather conditions make it more likely to see a delay?

```{r}
flights %>%
  group_by(origin,year,month,day) %>%
  mutate(avg_delay = mean(arr_delay, na.rm = T)) %>%
  select(origin,avg_delay,year,month,day) %>%
  inner_join(weather) %>%
  arrange(-avg_delay)
```

Cold weather, high winds, and low visibility

5. What happened on June 13 2013? Display the spatial pattern of delays, and then use Google to cross-reference with the weather

```{r}
flights5 <- flights %>%
  filter(month == 6 & day == 13) %>%
  group_by(dest) %>%
  mutate(avg_delay = mean(arr_delay, na.rm = T)) %>%
  select(dest,avg_delay) %>%
  inner_join(airports, by = c(dest = "faa"))

flights5 %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point(aes(color = avg_delay)) +
    coord_quickmap()
```

Something going down on east coast. Everyone on the east coast stopped to watch an Australian swim in Florida. Or more likely the June 12–13, 2013 derecho series. A widespread, long-lived, straight-line wind storm that is associated with a land-based, fast-moving group of severe thunderstorms.

### 13.4.7 Other implementations

## 13.5 Filtering joins

```{r}
top_dest <- flights %>%
  count(dest, sort = TRUE) %>%
  head(10)
top_dest

flights %>% 
  filter(dest %in% top_dest$dest)

flights %>% 
  semi_join(top_dest)

flights %>%
  anti_join(planes, by = "tailnum") %>%
  count(tailnum, sort = TRUE)
```

### 13.5.1 Exercises

1. What does it mean for a flight to have a missing `tailnum`? What do the tail numbers that don’t have a matching record in `planes` have in common? (Hint: one variable explains ~90% of the problems.)

```{r}
flights %>%
  anti_join(planes, by = "tailnum")
flights %>%
  anti_join(planes, by = "tailnum") %>%
  count(carrier, sort = T)
flights %>%
  filter(carrier == "AA") %>%
  count(tailnum)
flights %>%
  filter(carrier == "MQ") %>%
  count(tailnum)
```
AA and MQ did not report a significant number of their tailnums

2. Filter flights to only show flights with planes that have flown at least 100 flights

```{r}
enough <- flights %>%
  group_by(tailnum) %>%
  count(tailnum) %>%
  filter(n >= 100)

flights %>%
  semi_join(enough, by = "tailnum")
```

3. Combine `fueleconomy::vehicles` and `fueleconomy::common` to find only the records for the most common models.

```{r}
fueleconomy::vehicles
fueleconomy::common
fueleconomy::vehicles %>%
  semi_join(fueleconomy::common, by = "model")
```

4. Find the 48 hours (over the course of the whole year) that have the worst delays. Cross-reference it with the `weather data`. Can you see any patterns?
```{r}
delays <- flights %>%
  group_by(month,day) %>%
  mutate(delay = mean(arr_delay, na.rm = T)) %>%
  select(month,day,delay) %>%
  unique() %>%
  arrange(month,day)
two_day_delay <- roll_sum(delays$delay, n = 2)
grep(max(two_day_delay), two_day_delay)
worst <- delays[203:204,]
worst
inner_join(worst,weather) %>%
  select(-delay)
```
July 22nd and 23rd had the worst combined delays

5. What does `anti_join(flights, airports, by = c("dest" = "faa"))` tell you? What does `anti_join(airports, flights, by = c("faa" = "dest"))` tell you?

```{r}
anti_join(flights, airports, by = c("dest" = "faa"))
anti_join(airports, flights, by = c("faa" = "dest"))
```

The first shows all the flights whose destinations are not contained in the airport data frame's faa column. The second shows all airports which have a value in the faa column which is not present in the flights dest column.

6. You might expect that there’s an implicit relationship between plane and airline, because each plane is flown by a single airline. Confirm or reject this hypothesis using the tools you’ve learned above.

```{r}
flights %>%
  select(carrier,tailnum) %>%
  unique() %>%
  count(tailnum) %>%
  filter(n > 1)
```

Not true, there are at least 17 planes which were flown under 2 different airlines in 2013