---
title: "05_23_2017"
author: "John D."
date: "May 23, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(reshape2)
```

# 7.3 Variation

### 7.3.1 Visualising distributions

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))

diamonds %>% 
  count(cut)

ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = carat), binwidth = 0.5)

diamonds %>% 
  count(cut_width(carat, 0.5))

smaller <- diamonds %>% 
  filter(carat < 3)

ggplot(data = smaller, mapping = aes(x = carat)) +
  geom_histogram(binwidth = 0.1)

ggplot(data = smaller, mapping = aes(x = carat, colour = cut)) +
  geom_freqpoly(binwidth = 0.1)
```

### 7.3.2 Typical values

```{r}
ggplot(data = smaller, mapping = aes(x = carat)) +
  geom_histogram(binwidth = 0.01)

ggplot(data = faithful, mapping = aes(x = eruptions)) + 
  geom_histogram(binwidth = 0.25)

```

### 7.3.3 Unusual values

```{r}
ggplot(diamonds) + 
  geom_histogram(mapping = aes(x = y), binwidth = 0.5)

ggplot(diamonds) + 
  geom_histogram(mapping = aes(x = y), binwidth = 0.5) +
  coord_cartesian(ylim = c(0, 50))

unusual <- diamonds %>% 
  filter(y < 3 | y > 20) %>% 
  select(price, x, y, z) %>%
  arrange(y)
unusual

```

### 7.3.4 Exercises

1. Explore the distribution of each of the `x`, `y`, and `z` variables in `diamonds`. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.
```{r}
summary(diamonds)
smaller <- diamonds %>% select(x,y,z)
colMeans(smaller)
m.smaller <- melt(smaller)
ggplot(m.smaller) + 
  geom_histogram(aes(value)) + 
  facet_wrap(~variable) +
  coord_cartesian(ylim = c(0, 50))
```

Based on the distribution, I believe x and y are the width and height and z is the depth. The average of x and y are very similar which could be due to people not wanting oblong diamonds. Additionally, x and y are on average larger than z, suggesting people want diamonds that cover a lot of surface area but do not necessarily care about the depth.

2. Explore the distribution of `price`. Do you discover anything unusual or surprising? (Hint: Carefully think about the binwidth and make sure you try a wide range of values.)

```{r}
smaller <- diamonds %>% select(price,cut,clarity)
summary(smaller)

p <- ggplot(smaller, aes(x = price))
p + geom_histogram()
p + geom_histogram(binwidth = 10)
p + geom_histogram(binwidth = 100)
p + geom_histogram(binwidth = 1000)
p + geom_histogram(binwidth = 2000)
p + geom_histogram(binwidth = 10)
p + geom_histogram(binwidth = 10) + coord_cartesian(xlim = c(1400,1600))
```

There seems to be no diamonds priced between $1460 and $1540

3. How many diamonds are 0.99 carat? How many are 1 carat? What do you think is the cause of the difference?

```{r}
diamonds %>% 
  filter(carat == .99 | carat == 1) %>%
  group_by(carat) %>%
  count()

```

There are 1558 1 carat diamonds and 23 .99 carat diamonds. This is most likely due to people wanting their diamonds to be atleast 1 carat big. That or people like diamonds with carats that are a multiple of .25, .99 is an unusual size for a cut diamond.

4. Compare and contrast `coord_cartesian()` vs `xlim()` or `ylim()` when zooming in on a histogram. What happens if you leave binwidth unset? What happens if you try and zoom so only half a bar shows?

```{r}
smaller <- diamonds %>% select(price)
p <- ggplot(smaller, aes(x = price))

p + geom_histogram(binwidth = 100) + coord_cartesian(xlim = c(0,1000))
p + geom_histogram(binwidth = 10000) + coord_cartesian(xlim = c(0,1000))
p + geom_histogram(binwidth = 100) + xlim(0,1000)

p + geom_histogram() + coord_cartesian(xlim = c(0,1000))
p + geom_histogram() + xlim(0,1000)
```

`xlim()` or `ylim()` will replace values outside of the range with NA and will not be included when creating the histogram and determining bin width. `coord_cartesian()` simply zooms in on the certain part of the plot without changing the bin width.

# 7.4 Missing values

```{r}
diamonds2 <- diamonds %>% 
  filter(between(y, 3, 20))

diamonds2 <- diamonds %>% 
  mutate(y = ifelse(y < 3 | y > 20, NA, y))

ggplot(data = diamonds2, mapping = aes(x = x, y = y)) + 
  geom_point()

ggplot(data = diamonds2, mapping = aes(x = x, y = y)) + 
  geom_point(na.rm = TRUE)

nycflights13::flights %>% 
  mutate(
    cancelled = is.na(dep_time),
    sched_hour = sched_dep_time %/% 100,
    sched_min = sched_dep_time %% 100,
    sched_dep_time = sched_hour + sched_min / 60
  ) %>% 
  ggplot(mapping = aes(sched_dep_time)) + 
    geom_freqpoly(mapping = aes(colour = cancelled), binwidth = 1/4)
```

### 7.4.1 Exercises

1. What happens to missing values in a histogram? What happens to missing values in a bar chart? Why is there a difference?

```{r}
diamonds2 <- diamonds %>% 
  mutate(price = ifelse(price < 500 | price > 20000, NA, price))
ggplot(data = diamonds2, mapping = aes(x = price)) + 
  geom_histogram()
diamonds2 <- diamonds %>% 
  mutate(cut = ifelse(price < 500 | price > 20000, NA, as.character(cut)))
ggplot(data = diamonds2, mapping = aes(x = cut)) + 
  geom_bar()
```

In a histogram the NA values are removed and you are warned. In a bar plot, the NA values get their own bar. Has to do with the difference between categorical and continous variables.

2. What does `na.rm = TRUE` do in `mean()` and `sum()`?

```{r}
diamonds2 <- diamonds %>% 
  mutate(price = ifelse(price < 500 | price > 20000, NA, price))
mean(diamonds2$price)
mean(diamonds2$price, na.rm = T)
sum(diamonds2$price)
sum(diamonds2$price, na.rm = T)
```

If `na.rm` is `FALSE` an `NA` or `NaN` value in any of the arguments will cause a value of `NA` or `NaN` to be returned, otherwise `NA` and `NaN` values are ignored. `na.rm = TRUE` makes it so NAs are not included in the calculations

# 7.5 Covariation

```{r}
ggplot(data = diamonds, mapping = aes(x = price)) + 
  geom_freqpoly(mapping = aes(colour = cut), binwidth = 500)

ggplot(diamonds) + 
  geom_bar(mapping = aes(x = cut))

ggplot(data = diamonds, mapping = aes(x = price, y = ..density..)) + 
  geom_freqpoly(mapping = aes(colour = cut), binwidth = 500)

ggplot(data = diamonds, mapping = aes(x = cut, y = price)) +
  geom_boxplot()

ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
  geom_boxplot()
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = reorder(class, hwy, FUN = median), y = hwy))
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = reorder(class, hwy, FUN = median), y = hwy)) +
  coord_flip()
```

### 7.5.1.1 Exercises

1. Use what you’ve learned to improve the visualisation of the departure times of cancelled vs. non-cancelled flights.

```{r}
df <- nycflights13::flights %>% 
  mutate(
    cancelled = is.na(dep_time),
    sched_hour = sched_dep_time %/% 100,
    sched_min = sched_dep_time %% 100,
    sched_dep_time = sched_hour + sched_min / 60
  )

ggplot(df, mapping = aes(sched_dep_time)) + 
  geom_density(mapping = aes(colour = cancelled))
ggplot(df, mapping = aes(y = sched_dep_time)) + 
  geom_boxplot(mapping = aes(x = cancelled))
```

2. What variable in the diamonds dataset is most important for predicting the price of a diamond? How is that variable correlated with cut? Why does the combination of those two relationships lead to lower quality diamonds being more expensive?

```{r}
summary(diamonds)
n.diamonds <- diamonds
n.diamonds$cut <- as.integer(n.diamonds$cut)
n.diamonds$clarity <- as.integer(n.diamonds$clarity)
n.diamonds$color <- as.integer(n.diamonds$color)
correlations <- cor(n.diamonds)
correlations

ggplot(diamonds, aes(cut,price)) + geom_boxplot()
ggplot(diamonds, aes(carat,price)) + geom_point(aes(color = cut))
```

Carat is most important for for predicting the price of a diamond. As the carat increases, the price rapidly increases regardless of cut. This is what causes lower quality diamonds to be more expensive

3. Install the ggstance package, and create a horizontal boxplot. How does this compare to using coord_flip()?

```{r}
#install.packages("ggstance")
library(ggstance)

ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = reorder(class, hwy, FUN = median), y = hwy)) +
  coord_flip()

ggplot(data = mpg) +
  geom_boxploth(mapping = aes(x = reorder(class, hwy, FUN = median), y = hwy))

ggplot(data = mpg) +
  geom_boxploth(mapping = aes(y = reorder(class, hwy, FUN = median), x = hwy))
```

Creates the same horinzontal boxplot, need to switch the x and y variables though.

4. One problem with boxplots is that they were developed in an era of much smaller datasets and tend to display a prohibitively large number of “outlying values”. One approach to remedy this problem is the letter value plot. Install the lvplot package, and try using `geom_lv()` to display the distribution of price vs cut. What do you learn? How do you interpret the plots?

```{r}
#install.packages("lvplot")
library(lvplot)
ggplot(data = diamonds) +
  geom_lv(aes(cut,price))
```

An extension of standard boxplots which draws k letter statistics

5. Compare and contrast `geom_violin()` with a facetted `geom_histogram()`, or a coloured `geom_freqpoly()`. What are the pros and cons of each method?

```{r}
ggplot(data = diamonds) +
  geom_violin(aes(cut,price))
ggplot(data = diamonds) +
  geom_histogram(aes(price)) +
  facet_wrap(~cut)
ggplot(data = diamonds, mapping = aes(x = price, y = ..density..)) + 
  geom_freqpoly(mapping = aes(colour = cut), binwidth = 500)
```

A violin plot is a like a faceted histgram except all the x variables are on the same plot. a frequency plot is similar except they are all overlapping each other.

6. If you have a small dataset, it’s sometimes useful to use `geom_jitter()` to see the relationship between a continuous and categorical variable. The ggbeeswarm package provides a number of methods similar to `geom_jitter()`. List them and briefly describe what each one does.

```{r}
#install.packages("ggbeeswarm")
library(ggbeeswarm)
qplot(class, hwy, data = ggplot2::mpg, geom='beeswarm')

```

geom_beeswarm creates a plot similar to a violin plot but with dots.
The beeswarm geom is a convenient means to offset points within categories to reduce overplotting.

```{r}
ggplot(data = diamonds) +
  geom_count(mapping = aes(x = cut, y = color))

diamonds %>% 
  count(color, cut)

diamonds %>% 
  count(color, cut) %>%  
  ggplot(mapping = aes(x = color, y = cut)) +
    geom_tile(mapping = aes(fill = n))
```

###7.5.2.1 Exercises

1. How could you rescale the count dataset above to more clearly show the distribution of cut within colour, or colour within cut?

```{r}
diamonds %>% 
  count(color, cut) %>%  
  group_by(color) %>%
  mutate(distrib = n / sum(n)) %>%
  ggplot(mapping = aes(x = color, y = cut)) +
    geom_tile(mapping = aes(fill = distrib)) +
  scale_fill_gradient2(limits = c(0,.5))

diamonds %>% 
  count(color, cut) %>%  
  group_by(cut) %>%
  mutate(distrib = n / sum(n)) %>%
  ggplot(mapping = aes(x = color, y = cut)) +
    geom_tile(mapping = aes(fill = distrib)) +
  scale_fill_gradient2(limits = c(0,.5))

```

2. Use `geom_tile()` together with dplyr to explore how average flight delays vary by destination and month of year. What makes the plot difficult to read? How could you improve it?

```{r}
library(nycflights13)
flights %>% 
  group_by(dest,month) %>%
  mutate(avg_delay = mean(arr_delay, na.rm = T)) %>%
  select(dest,month,avg_delay) %>%
  unique() %>%
  ungroup() %>%
  ggplot(mapping = aes(x = month, y = dest)) +
  geom_tile(mapping = aes(fill = avg_delay))

```

Theres a lot of possible combinations, 1113 possible. Can improve it by condensing the data into different groups like region or quarter of month.

3. Why is it slightly better to use `aes(x = color, y = cut)` rather than `aes(x = cut, y = color)` in the example above?

```{r}
diamonds %>% 
  count(color, cut) %>%  
  ggplot(mapping = aes(x = color, y = cut)) +
    geom_tile(mapping = aes(fill = n))

diamonds %>% 
  count(color, cut) %>%  
  ggplot(mapping = aes(y = color, x = cut)) +
    geom_tile(mapping = aes(fill = n))

```

It's a little bit more visually pleasing to have the larger categorical variable on the x axis. Also the x axis names aren't smooshed in the viewer.

```{r}
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = carat, y = price))

ggplot(data = diamonds) + 
  geom_point(mapping = aes(x = carat, y = price), alpha = 1 / 100)

#install.packages("hexbin")
library(hexbin)

smaller <- diamonds %>% 
  filter(carat < 3)

ggplot(data = smaller) +
  geom_bin2d(mapping = aes(x = carat, y = price))
ggplot(data = smaller) +
  geom_hex(mapping = aes(x = carat, y = price))

ggplot(data = smaller, mapping = aes(x = carat, y = price)) + 
  geom_boxplot(mapping = aes(group = cut_width(carat, 0.1)))

ggplot(data = smaller, mapping = aes(x = carat, y = price)) + 
  geom_boxplot(mapping = aes(group = cut_width(carat, 0.1)), varwidth = T)

ggplot(data = smaller, mapping = aes(x = carat, y = price)) + 
  geom_boxplot(mapping = aes(group = cut_number(carat, 20)))
```

### 7.5.3.1 Exercises

1. Instead of summarising the conditional distribution with a boxplot, you could use a frequency polygon. What do you need to consider when using `cut_width()` vs `cut_number()`? How does that impact a visualisation of the 2d distribution of `carat` and `price`?

When using `cut_width()` and `cut_number()` you need to consider the distribution of your data. As seen above, `cut_number(carat, 20)` creates boxplots with a large distribution in size. With `cut_width()` you may end up grouping data that aren't very similar but get binned together due to the cut offs and get a larger distribution of number of points in each boxplot. Improper values for both will reduce your ability to interpret your data.

2. Visualise the distribution of carat, partitioned by price.

```{r}
ggplot(data = smaller, mapping = aes(x = price, y = carat)) + 
  geom_boxplot(mapping = aes(group = cut_width(price, 1000)), varwidth = T)

```

3. How does the price distribution of very large diamonds compare to small diamonds. Is it as you expect, or does it surprise you?

```{r}
summary(diamonds$carat) # median == .7

smaller <- diamonds %>% 
  filter(carat <= .7)
larger <- diamonds %>% 
  filter(carat > .7)

ggplot(data = smaller) +
  geom_bin2d(mapping = aes(x = carat, y = price))

ggplot(data = larger) +
  geom_bin2d(mapping = aes(x = carat, y = price))

ggplot(data = diamonds) +
  geom_bin2d(mapping = aes(x = carat, y = price))
```

Smaller diamonds' price grows at a linear rate and the prices have a nice distribution. After you pass 1 carat though the distribution goes wild and price ranges increase exponentially.

4. Combine two of the techniques you’ve learned to visualise the combined distribution of cut, carat, and price.

```{r}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_boxplot(aes(fill = cut, cut_width(carat, 1)))
```

5. Two dimensional plots reveal outliers that are not visible in one dimensional plots. For example, some points in the plot below have an unusual combination of `x` and `y` values, which makes the points outliers even though their `x` and `y` values appear normal when examined separately. Why is a scatterplot a better display than a binned plot for this case?

```{r}
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = x, y = y)) +
  coord_cartesian(xlim = c(4, 11), ylim = c(4, 11))

ggplot(data = diamonds) +
  geom_boxplot(mapping = aes(x = x, y = y, cut_number(x, 30))) +
  coord_cartesian(xlim = c(4, 11), ylim = c(4, 11))
```

When you bin, you have the chance of masking and/or adding outliers. Outliers are much easier to see in the first plot than in the second plot. Also outliers in the second plot may simply be a result of inappropriate binning. Also with the plots above, the data is much easier to see in the first plot than in the second plot. 